### 2010-05-19 paper/paper.tex
brief outline from amanda's email

\documentclass[12pt]{report}
\usepackage[pdftex]{graphicx}
\usepackage{url}
\usepackage{amsmath}
\usepackage{subfig}
\usepackage{listings}
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
\hyphenation{auto-nomous}

\begin{document}
\title{Rutgers Autonomous Aircraft Team\\Technical Report\\2010 AUVSI UAS Competition}
\author{Patrick Hickey et. al.}
\input{./title.tex}

\begin{abstract}
we will put the abstract here.
\end{abstract}

\section{Introduction}

Brief overview of what we accomplished.

\section{Systems Engineering}

\section{Flight Vehicle}

\subsection{Airframe}

\subsection{Power}

\subsubsection{Engine}

\subsubsection{Batteries}

\section{Autopilot}

\subsection{Requirments}

\subsection{Hardware Design}

\subsection{Software Design}

\subsection{Ground Station}

\subsection{Tuning and Testing}

\section{Imaging System}

\subsection{Camera}

\subsection{Image Processing}

\subsection{Ground Station Interface}

\section{Operation Safety}

\subsection{Safety Features}

\subsection{Procedures}

\end{document}

### 2010-05-19 paper/title.tex
title page created

\begin{titlepage}
\begin{center}

\includegraphics[width=0.5\textwidth]{../images/daedalus.jpg}\\[1cm]
\HRule \\[1cm]
{ \huge \bfseries Rutgers University Autonomous Aircraft Team } \\[0.5cm]
\HRule \\[0.5cm]
{ \large \bfseries Technical Report }
\\[0.5cm]
{ \large \bfseries 2010 AUVSI UAS Competition }
\\[0.5cm]
\HRule \\[1cm]

  {\large Amanda Gaetano, Anthony Garrison, Pat Hickey,}
\\{\large Stephen Indyk, Cogan Noll, John Palmer,}
\\{\large Adrien Perkins, Gregory Quinn, and Michael Varga}
\\[1cm]
\emph{Thank you to our sponsors}
\\ Rutgers University Engineering Governance Council
\\ Rutgers Alumni Association
\\ WINLAB
\\ Invensense, Inc.
\\ ST Micro, Inc.

\vfill
{\large \today}

\end{center}
\end{titlepage}

### 2010-05-19 paper/paper.tex
Added text to airframe section, still needs work

@@
-\documentclass[12pt]{report}
+\documentclass[10pt]{report}
 \usepackage[pdftex]{graphicx}
-\usepackage{url}
+\usepackage{hyperref}
+\hypersetup{ colorlinks=true
+           , linkcolor=cyan
+           , pdfnewwindow=true 
+           }
 \usepackage{amsmath}
@@
 \hyphenation{auto-nomous}
+
+% Limited by competition rules to min font size 10 pt, max pages 20
 
@@
 \subsection{Airframe}
+The Rutgers team, after much deliberation, elected to persue a kit-type airplane rather than an in-house design.
+This decision was based on a thorough analysis of available time and materials: though an in-house design is always preferred, it was quickly determined that the team, in its second year, did not yet posses the time or expertise required for a project of such enormity.
+Eager to avoid mistakes made in the previous year that resulted in a catastrophic crash mere weeks before  competition, the team gladly accepted an already-constructed plane from a local RC club.
+The donating member intimated that the model was in good, ready-to-fly condition, requiring only minimal repairs.
+Upon further inspection, the team readily came to the conclusion that, unfortunately, a significant remodel was necessary: wood was rotted out, foam eaten away, and necessary structural members non-existant.
 
@@
 \section{Autopilot}
+Our entry uses an autopilot system based off the open source 
+Paparazzi project\cite{paparazziweb}. 
+We ported the airborne code to Linux in order to use a single computer for all of our flight hardware and software.
 
@@
 
-\subsection{Hardware Design}
+We needed to interface with the following hardware components:
+\begin{itemize}
+	\item GPS
+	\item servos
+	\item etc.
+\end{itemize}
 
-\subsection{Software Design}
+\subsection{Hardware}
+
+\subsection{Software}
 
@@
 
+\bibliographystyle{plain}
+\bibliography{paper}
+
 \end{document}

### 2010-05-19 paper/paper.tex
filled in a lot of the outline with contributor responsibilities

@@
 \begin{abstract}
-we will put the abstract here.
+We will write the abstract after the rest of the paper is finished. If we're really close to the page limit, we may be able to omit it.
 \end{abstract}
@@
 
-Brief overview of what we accomplished.
+Brief overview of what we accomplished. This should be a very brief chronology of the engineering we've done to date.
 
@@
 
+We need to give an overview of our airframe, autopilot, and payload requirments. Amanda or Anthony should decide how to proceed here based on other good papers.
+
+One thing we need is a good justification for having built two fucking giant airframes.
+Lets point out that we initially assumed a much heavier payload, but found lighter and smaller alternatives as the year went on.
+We decided it was better to be too big than too small.
+A large plane with lower wing loading would be easier to control, and easier to recover in case power is lost.
+(someone check that math)
+
 \section{Flight Vehicle}
 
+At the time of this writing, we have completed and flown a single flight vehicle, which we have named the \emph{Daedalus}.
+We have also nearly completed a very similar backup airframe which serve as a backup in case the \emph{Daedalus} has a disaster. The \emph{Knight Two} will be fitted with
+
 \subsection{Airframe}
+
+We elected to persue a kit-type airplane rather than our own design. 
+Our own custom design, the \emph{Icarus}, suffered a catastrophic structural failure last year.
+From this, we learned that custom structures require much more testing and revision than a proven kit.
+We also reconsidered our payload requirments and decided we could easily modify a kit plane to accept our autopilot and imaging system.
+
+\subsubsection{Daedalus}
+
+A member of our local R/C club, Tri-County RC \cite{tricountyRC}, donated a ten-foot wingspan high wing trainer to our club.
+The plane, first built out of foam and balsa from long-lost plans in 1986, was accepted graciously, but required quite a bit of work before it was once again flightworthy.
+We removed the covering to find water damage and rot which resulted from years of storage.
+We stripped and rebuilt nearly the entire airframe, adding carbon-fiber reinforcements to the wing in anticipation of increased wing loading due to our payload.
+
+Once rebuilt and covered, we tested the \emph{Daedalus} in our shop to ensure the rebuilt structure was strong enough.
+Amanda and Anthony, please continue working on this paragraph, describing our testing procedures
+
+Figure goes here: Photos of Daedalus in construction, sandbag testing
+
+Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations
+
+Figure may go here: Line drawing of Daedalus. JOHN
+
+\subsubsection{Knight Two}
+
+In order to mitigate a disaster such as we had last year, we decided to build another airframe which could replace the \emph{Daedalus} at short notice.
+We selected an Aero-Design 12 Foot Telemaster \cite{aerodesign} because it was available as a balsa kit and similar in size to the \emph{Daedalus}.
+
+We named it the \emph{Knight Two} after the Rutgers University mascot, the Scarlet Knight.
+We may rename it if someone can think of a better name.
+Pat needed to think up a decent name real fast in order to make this rough draft.
+
+Since our testing plan for the \emph{Daedalus} was successful, we will follow the same plan when, in the coming weeks, we finish building \emph{Knight Two}.
+
+Figure goes here: Photos of Knight Two in construction, sandbag testing
+
+Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations
+
+Figure may go here: Line drawing of Knight Two. JOHN
 
@@
 
+Each airframe has a 45cc gasoline engine.
+Each airframe has a xx ounce gasoline tank.
+Based on our flight tests, with climbs to 400 feet AGL, a full xx ounce gasoline tank gives us xx minutes of flight time.
+(can someone guestimate this?
+Do we have data from break-in, etc)
+
 \subsubsection{Batteries}
+
+Each airframe has nearly identical flight electronics.
+To power a Futaba 2.4GHz receiver and 6 servos for a XX minute flight, we selected a 2 cell XX amp hour Lithium Ion battery with a 5v, 10A capacity switching regulator made by Castle Creations.
+Pat will add more here if he has to.
 
@@
 	\item servos
+	\item camera
 	\item etc.
@@
 
+Interfacing with such a large array of devices meant we had to use a ...
+
 \subsection{Hardware}
+Single board computer: beagleboard.
 
+Description of each input, output interface.
+
+Figure goes here: Hardware connection diagram. PAT
 \subsection{Software}
+PAT
 
+Figure goes here: interprocess communication. PAT
 \subsection{Ground Station}
+Because we based our autopilot software on the Paparazzi project, we were able to make use of their excellent ground control software (GCS) with essentially no modifications.
 
+Detail the capabilities of the Paparazzi GCS.
+
+Figure goes here: Paparazzi ground station screenshot, or whatever. 
 \subsection{Tuning and Testing}
+PAT
 
 \section{Imaging System}
+COGAN, MICHAEL write an overview
 
@@
 
+MICHAEL : physical specification of camera (size, 5 megapixel stills). We selected it because...
+
+MICHAEL: camera interfaces: serial and USB
+
+The camera will be mounted in a pan and tilt unit, which will ... STEVE, PAT
+
 \subsection{Image Processing}
+MICHAEL, describe flow of control and images in the camera software.
+
+COGAN, describe computer vision algorithms, and anything else we do
 
@@
 
+COGAN, describe the ground station interface
+
+Figure goes here: COGAN, provide a screen shot of the ground station in 'action.' We should try to use images that at least look like they were taken of targets from above.
+Maybe take pictures from 7th floor CORE, like we said we were going to use to train the cv algorithms.
+
 \section{Operation Safety}
+
+PAT
 
@@
 
+PAT
+
 \subsection{Procedures}
+
+PAT (well i dont know what to do here yet)
 

### 2010-05-19 paper/paper.tex
added info about Daedalus, started to outline operational safety

@@
 Once rebuilt and covered, we tested the \emph{Daedalus} in our shop to ensure the rebuilt structure was strong enough.
-Amanda and Anthony, please continue working on this paragraph, describing our testing procedures
+Of particular concern was interface between the wings and the center section---it was important to confirm that the wings were mounted and reinforced in a manner that would permit them to carry the load of the plane while in flight.
+To this end, we put the \emph{Daedalus} through a series of static and dynamic ``sandbag tests.''  The plane was carefully supported under the center section, nose, and tail while pre-weighted bags of sand were slowly placed symetrically and simultaneously on the underside of each wing, starting from the innermost rib.
+The \emph{Daedalus} easily supported XX lbs (XX percent of its weight) under static loading.
+We then lifted the plane off its supports and jostled it to simulate in-flight motion.
+The plane seemed to remain stable, and upon further observation after unloading, no cracks or defects were detected.
+
+Satisfied, we proceeded through the usual engine and surface tests before taxiing the plane around a parking lot.
+We brought the plane back inside, examined it for signs of stress and, finding everything in good condition, deemed the \emph{Daedalus} ready to fly.
 
@@
 
-PAT
+We consider the safety of our team, our aircraft, and our spectators to be of utmost importance.
+As such, we've implemented numerous safety features and developed a strict procedure for starting, taxiing, and flying the plane.
+This procedure enables us to ensure that all personnel are safely in place, and that no intergral checks are missed.
 
 \subsection{Safety Features}
-
-PAT
+Our system is designed to gracefully handle any failure while continuing to ensure the safety of those nearby.
+\begin{itemize}
+\item Channel 5 (?) allows quick switch to safety pilot in case of autopilot malfunction
+\item autopilot return home command if connection is lost (I presume it does something like that?)
+\item kill switch
+\item etc?
+\end{itemize}
 
 \subsection{Procedures}
-
-PAT (well i dont know what to do here yet)
+Our team employs a clearly defined flight routine every time the engine is started.
+Blah blah blah.
+Maybe just include either the flight routine or pre-flight checklist as an appendix?
+I feel like this stuff is pretty standard, and there's probably no point in us wasting a lot of space rambling about it.
 

### 2010-05-21 paper/paper.tex
fixed paragraph breaks
systems engineering from adrien-needs work, will fix tonight

@@
 
-We need to give an overview of our airframe, autopilot, and payload requirments. Amanda or Anthony should decide how to proceed here based on other good papers.
+The main systems engineering approach that was used in order to finish construction of the planes efficiently and in a timely manner was to break the team into two main groups: the mechanical team and the electronics team.
+The mechanical team was in charge of the rebuild and build process of the \emph{Daedalus} and the \emph{Knight Two}, respectively.
+In order to maximize efficiency we further divided the mechanical team into sub groups placing members where they were best suited.
+During both constructions we split into a wing and tail section group and a fuselage group.
+While all the construction was going on, two expert builders from the local R/C club would come and advise us when we were unsure how to proceed in order to have the best balance of weight and strength.
+Along with the actual construction of the planes, the mechanical team also focused on modifying the structures of the planes in order to fit the desired payload: the autopilot system and the imaging system.
 
-One thing we need is a good justification for having built two fucking giant airframes.
+The electronics team was in charge of the autopilot system and the imaging system.
+Originally a large and heavy system was to be used for an imaging system, but as the year progress a lighter and smaller alternative was found, which helped the mechanical team in the placement of the imaging system…(I do not know the details of what they did, that’s all Pat)
+
+We will be performing a manual take off, and giving the autopilot system control of the aircraft once the plane is in the air.
+For landing, we will do the same process in reverse, and land the plane manually.
+
+%One thing we need is a good justification for having built two fucking giant airframes.
 Lets point out that we initially assumed a much heavier payload, but found lighter and smaller alternatives as the year went on.

### 2010-05-22 paper/paper.tex
cogan's additions on image station

--- 2010-05-19 paper/paper.tex
@@
 \usepackage{listings}
+\usepackage{float}
 \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
@@
 \section{Imaging System}
-COGAN, MICHAEL write an overview
+
+To obtain images of targets in flight we have created an Imaging System capable of meeting all the requirements of the competition.
+The system consists of a pan-tilt camera on the plane, communication software on the plane, and our ``Image Station'' application running on a dedicated laptop.
+The Image Station communicates with the plane in real time via two different sets of wireless radios.
+Using the Image Station, the operator can view a live video stream from the plane, control the pan-tilt camera remotely, and download and manipulate images.
 
@@
 
-\subsection{Image Processing}
-MICHAEL, describe flow of control and images in the camera software.
+\subsection{Obtaining Images}
 
-COGAN, describe computer vision algorithms, and anything else we do
+The entire imaging system is controlled by the ``Image Station Operator''.
+They view a live video stream coming from the camera on the plane, which is transmitted over a high bandwidth radio.
+The operator can also control the camera by sending commands to the plane over an Xbee Radio link.
+In order to assist the operator in finding targets, a target locating algorithm will be run on the video feed to highlight possible targets.
+At the time of writing work is in progress on this feature, and it is being implemented using the OpenCV library \cite{opencv} with contour detection and polygonal mapping functions.
 
-\subsection{Ground Station Interface}
+\begin{figure} [H]
+  \centering
+  	\includegraphics[width=0.9\textwidth]{../images/ImageStationControls.jpg}
+  	\caption[Image Station Control Panel]{The Control Panel for the Image Station.
+Above is a live video feed from the plane.
+Below are controls to adjust the pan/tilt of the unit, zoom in/out, take a picture, and reset the camera to a neutral position.}
+  	\label{imagestationcontrols}
+\end{figure}
 
-COGAN, describe the ground station interface
+When a picture is taken on the plane it is immediately saved on the camera's memory card.
+In order to view that picture on the ground it must first be transferred to the plane's main computer (beagle board), and then downloaded to the Image Station over the Xbee connection.
 
-Figure goes here: COGAN, provide a screen shot of the ground station in 'action.' We should try to use images that at least look like they were taken of targets from above.
-Maybe take pictures from 7th floor CORE, like we said we were going to use to train the cv algorithms.
+\subsection{Downloading Images in Real Time}
+
+Once images are on the plane's computer they are ready to be downloaded over the Xbee Radio link.
+Each image is fairly large, and our bandwidth is only around 115200 baud, so we use a bandwidth conserving method to transfer images.
+The entire image is first downloaded as a low-res `thumbnail.'  Once that is available for viewing, the operator can select an area of interest and download a high-res crop of just that area.
+This allows us to quickly obtain high resolution images of the targets of interest.
+
+\begin{figure} [h]
+  \centering
+  	\includegraphics[width=0.9\textwidth]{../images/ImageStationMain.jpg}
+  	\caption[Image Station Interface]{The main interface for the Image Station.
+All the images that have been taken are displayed on the top left.
+The queue of images to download is on the bottom left.
+On the right is the currently displayed image, with target information underneath.}
+  	\label{imagestationinterface}
+\end{figure}
+
+\subsection{Identifying target characteristics}
+
+When an image has been fully downloaded to the ground targets can be identified and tagged with metadata.
+Information about the plane's orientation, gps coordinates, altitude, yaw/pitch/roll, and pan/tilt of the camera are sent down with the image.
+Using this information the GPS location of the target can be determined using the intrinsic properties of the camera and some basic trigonometry.
+target shape, color, alphanumeric, alphanumeric color, and orienttion are all entered manually by the operator.
 

### 2010-05-22 paper/paper.tex
when i imported cogan's work i deleted amanda's by accident. my bad

@@
 \section{Systems Engineering}
+The main systems engineering approach that was used in order to finish construction of the planes efficiently and in a timely manner was to break the team into two main groups: the mechanical team and the electronics team.
+The mechanical team was in charge of the rebuild and build process of the \emph{Daedalus} and the \emph{Knight Two}, respectively.
+In order to maximize efficiency we further divided the mechanical team into sub groups placing members where they were best suited.
+During both constructions we split into a wing and tail section group and a fuselage group.
+While all the construction was going on, two expert builders from the local R/C club would come and advise us when we were unsure how to proceed in order to have the best balance of weight and strength.
+Along with the actual construction of the planes, the mechanical team also focused on modifying the structures of the planes in order to fit the desired payload: the autopilot system and the imaging system.
 
-We need to give an overview of our airframe, autopilot, and payload requirments. Amanda or Anthony should decide how to proceed here based on other good papers.
+The electronics team was in charge of the autopilot system and the imaging system.
+Originally a large and heavy system was to be used for an imaging system, but as the year progress a lighter and smaller alternative was found, which helped the mechanical team in the placement of the imaging system…(I do not know the details of what they did, that’s all Pat)
 
-One thing we need is a good justification for having built two fucking giant airframes.
+We will be performing a manual take off, and giving the autopilot system control of the aircraft once the plane is in the air.
+For landing, we will do the same process in reverse, and land the plane manually.
+
+%One thing we need is a good justification for having built two fucking giant airframes.
 Lets point out that we initially assumed a much heavier payload, but found lighter and smaller alternatives as the year went on.

### 2010-05-22 paper/paper.tex
added a lot of introduction, rewrote much of systems engineering

--- 2010-05-22 paper/paper.tex
@@
 
-Brief overview of what we accomplished. This should be a very brief chronology of the engineering we've done to date.
+The Rutgers Autonomous Aircraft Team was formed in September 2008 with the goal of entering the 2009 AUVSI UAS competition.
+We didn't make it to the 2009 competition after loosing our aircraft only four weeks before.
+Our imaging system and autopilot system were also far from ready.
+
+In terms of NCAA sport eligibility, one could consider 2009 our `red shirt' year. In 2010 we are still rookies.
+
+The team is made of N Rutgers undergraduates
+% List, can we do itemize for this? two columns? needs to be smaller
+Anthony Garrison, President
+Amanda Gaetano, Vice President
+Pat Hickey, Treasurer and Electronics Team Leader
+Greg Quinn, Mechanical Team Leader
+Stephen Indyk, Cogan Noll, John Palmer, Adrien Perkins, Michael Varga, Bradley Lord, David Wescott, and.....
+
+We are advised by
+ Professor Tobias Rossmann,
+ and several members of the Tri-County R/C Club \cite{tricountyRC}:
+ Larry Kosinar,
+ Bill (last name),
+ Mike Korb,
+ (the man who donated daedalus), and
+ Chris (test pilot) amongst many who spent time mentoring us.
+
+We have been generously funded by 
+Rutgers Engineering Governance Council,
+Rutgers Alumni Association, and
+WINLAB, a laboratory of the Rutgers University Department of Electrical Engineering.
+% What was Chris Brocco's Uncle's Company? Something Automation?
+BP Hobbies,
+
+We would like to acknowledge
+(Adrien's Father), for his prolific networking and promoting of the team;
+Gregg (last name) of BP Hobbies for going out of his way to help us;
+David Devries and Michael Maia, for their technical guidance;
+Professor Chung-chieh Shan, for overseeing Pat's independent study on the IMU design;
+Ivan Seskar, for keeping us (and many other undergrads) in the black;
+Roger Kondos and Brij Pathak, for contributing to the thermopile attitude sensor.
 
@@
 
-We need to give an overview of our airframe, autopilot, and payload requirments. Amanda or Anthony should decide how to proceed here based on other good papers.
+For this entry, we took a close look at our goals and requirments, and dialed back our approach to ensure we would meet the most basic competition requirments successfully.
+We started the design with some broad goals based on our experience in the 2009 competition. 
+Sub-teams and individuals tackled the smaller problems while still keeping an eye on the big picture.
 
-One thing we need is a good justification for having built two fucking giant airframes.
-Lets point out that we initially assumed a much heavier payload, but found lighter and smaller alternatives as the year went on.
-We decided it was better to be too big than too small.
-A large plane with lower wing loading would be easier to control, and easier to recover in case power is lost.
-(someone check that math)
+\subsubsection{Goals}
+\begin{itemize}
+	\item The air vehicle must be stable and slow, with ample payload capacity and 30 minute flight duration. It should be a proven and durable fixed-wing design.
+	\item The autopilot must be based on proven software, so we could focus on improving an existing system rather than reinventing the wheel.
+We felt the Paparazzi system was full featured and proven, and we wanted to adapt it to our needs.
+	\item The imaging system must be simple and capable.
+We assume the operator will be burdened with guiding the camera and identifying targets until time and resources permit us to add sensor information and control loops for guidance, and computer vision for identification.
+\end{itemize}
+
+We made design decisions of which competition goals we will attempt. 
+We will be performing a manual take off, and giving the autopilot system control of the aircraft once the plane is in the air.
+We will also perform a manual landing.
+We decided not to pursue the reach goals of autonomous takeoff and landing based on our inexperience and the risk, during development and competition, to our aircraft.
+
+One of the most difficult design decisions was to quantify ``ample payload.'' The payload size and weight is the major determining factor in the aircraft size.
+% -- expand this thought: 
+We started off assuming an electronics (computer + batteries) payload of 2.5lbs and a pan-tilt camera of 2.5lbs and 8 inch diameter.
+
+% Originally a large and heavy system was to be used for an imaging system, but as the year progress a lighter and smaller alternative was found, which helped the mechanical team in the placement of the imaging system(I do not know the details of what they did, that’s all Pat)
+
+Our team was divided by area of expertise into a mechanical team and an electrical team.
+The mechanical team began the year with the goal of producing two competition worthy airframes. These were to become the \emph{Daedalus} and \emph{Knight Two}.
+
+We determined, after consulting with R/C experts, that a high wing trainer is the most stable and proven fixed-wing design.
+Trainers traditionally have light wing loading for slow speed flight, and will be able to carry payload
+
+% -- donno how much use this is:
+%  During both constructions we split into a wing and tail section group and a fuselage group.
+While all the construction was going on, two expert builders from the local R/C club would come and advise us when we were unsure how to proceed in order to have the best balance of weight and strength.
+
+%Along with the actual construction of the planes, the mechanical team also focused on modifying the structures of the planes in order to fit the desired payload: the autopilot system and the imaging system.
+
+The electronics team was in charge of the autopilot system and the imaging system.
 

### 2010-05-22 paper/paper.tex
sm changes to introduction, systems engineering, itemize itemsep

@@
 We would like to acknowledge
-(Adrien's Father), for his prolific networking and promoting of the team;
-Gregg (last name) of BP Hobbies for going out of his way to help us;
-David Devries and Michael Maia, for their technical guidance;
+(Adrien's Father), for his prolific networking and promotion of the team;
+Gregg (last name) of BP Hobbies for getting us out of numerous binds;
+Dave Devries and Michael Maia, of Invensense, for their technical guidance;
 Professor Chung-chieh Shan, for overseeing Pat's independent study on the IMU design;
@@
 One of the most difficult design decisions was to quantify ``ample payload.'' The payload size and weight is the major determining factor in the aircraft size.
-% -- expand this thought: 
+% -- expand this thought pch: 
 We started off assuming an electronics (computer + batteries) payload of 2.5lbs and a pan-tilt camera of 2.5lbs and 8 inch diameter.
@@
 The electronics team was in charge of the autopilot system and the imaging system. 
+% not sure what i want to do to expand here pch
 
@@
 At the time of this writing, we have completed and flown a single flight vehicle, which we have named the \emph{Daedalus}.
-We have also nearly completed a very similar backup airframe which serve as a backup in case the \emph{Daedalus} has a disaster. The \emph{Knight Two} will be fitted with
+We have also nearly completed a very similar backup airframe which serve as a backup in case the \emph{Daedalus} has a disaster.
+The \emph{Knight Two} will be fitted with nearly identical R/C flight electronics, and accept the same autopilot and imaging system payload as the \emph{Daedalus}
 
@@
 
-Figure goes here: Photos of Knight Two in construction, sandbag testing
+Figure goes here: Photos of Knight Two in construction
 
@@
 \begin{itemize}
+	\setlength{\itemsep}{0cm}
+	\setlength{\parskip}{0cm}
 	\item GPS
@@
 Below are controls to adjust the pan/tilt of the unit, zoom in/out, take a picture, and reset the camera to a neutral position.}
-  	\label{imagestationcontrols}
+  	\label{fig:imagestationcontrols}
 \end{figure}
@@
 On the right is the currently displayed image, with target information underneath.}
-  	\label{imagestationinterface}
+  	\label{fig:imagestationinterface}
 \end{figure}
@@
 \begin{itemize}
+	\setlength{\itemsep}{0cm}
+	\setlength{\parskip}{0cm}
 	\item Channel 5 (?) allows quick switch to safety pilot in case of autopilot malfunction

### 2010-05-22 paper/paper.tex
added Taras to member list, Adriens father, edited operational safety

@@
 
-The team is made of N Rutgers undergraduates
+The team is made of 12 Rutgers undergraduates
 % List, can we do itemize for this? two columns? needs to be smaller
@@
 Greg Quinn, Mechanical Team Leader
-Stephen Indyk, Cogan Noll, John Palmer, Adrien Perkins, Michael Varga, Bradley Lord, David Wescott, and.....
+Stephen Indyk, Cogan Noll, John Palmer, Adrien Perkins, Michael Varga, Bradley Lord, David Wescott, and Taras Wallace.
 
@@
 We would like to acknowledge
-(Adrien's Father), for his prolific networking and promotion of the team;
+Jeff Perkins, for his prolific networking and promotion of the team;
 Gregg (last name) of BP Hobbies for getting us out of numerous binds;
@@
 Using this information the GPS location of the target can be determined using the intrinsic properties of the camera and some basic trigonometry.
-target shape, color, alphanumeric, alphanumeric color, and orienttion are all entered manually by the operator.
+Target shape, color, alphanumeric, alphanumeric color, and orientation are all entered manually by the operator.
 
@@
 	\setlength{\parskip}{0cm}
-	\item Channel 5 (?) allows quick switch to safety pilot in case of autopilot malfunction
-	\item autopilot return home command if connection is lost (I presume it does something like that?)
-	\item kill switch
+	\item Channel 5 (?) allows quick switch to safety pilot in case of autopilot malfunction.
+	\item If the plane loses contact with the ground station for more than XX seconds, the autopilot will default to a ``return home mode'' that will direct the plane towards the groundstation in hopes of reestablishing a connection.
+If the connection remains dead for XX minutes, the autopilot will direct the plane to [DO SOMETHING?]
+	\item Channel 6 (?) functions as a remote kill switch, ensuring the ability to cut power to the engine if an emergent situation precludes direct manual shut off.
 	\item etc?
@@
 Our team employs a clearly defined flight routine every time the engine is started.
-Blah blah blah.
-Maybe just include either the flight routine or pre-flight checklist as an appendix?
-I feel like this stuff is pretty standard, and there's probably no point in us wasting a lot of space rambling about it.
+This routine ensures that everyone in the starting area is fully alert and aware at all times, that no one stands in a position of known danger (i.e.
+the prop arc), and that all know when the starter makes contact with the spinner.
+A set script with clear key words not only ensures the safety of our team, but also is conducive to efficiency, consistency, and repeatability.
 

### 2010-05-22 paper/paper.tex
added some information to introduce steve's part of the report

@@
 \usepackage{float}
+\usepackage{textcomp}
 \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
@@
 
+\subsection{Pan Tilt Unit}
+\begin{figure}
+	\centering
+	\subfloat[Front View]{\label{fig:camera_frontview} 
+			\includegraphics[width=0.3\textwidth]{../images/camera_frontview.jpg}}
+	\subfloat[Isometric View]{\label{fig:camera_isometric} 
+			\includegraphics[width=0.3\textwidth]{../images/camera_isometric.jpg}}
+	\caption{Sony MHS-PM5. Image Credit: Sony}
+	\label{fig:camera}
+\end{figure}
+
+\begin{figure}
+
+\end{figure}
+
+In order to meet the target imaging requirments, we needed to add both pan and tilt actuators to the camera.
+The XX by YY degree lens field of view is insufficient to see off flight path targets, and would limit our capability in the target search area phase.
+We designed a pan tilt unit which make it possible to image the entire hemisphere below the fuselage of the aircraft.
+
+We took advantage of the MHS-PM5's swiveling lens assembley 
+(see Figure \ref{fig:camera_frontview}, and added a servo actuator which could rotate the lens 135\textdegree back from straight forward, as shown in Figure \ref{}
+The camera and tilt servo assembley is attached to the aircraft on a rotating mount.
+
 \subsection{Obtaining Images}

### 2010-05-22 paper/paper.tex
moved camera pictures figure. added pantilt figures
fixed merge with amanda's multicol commit

@@
 \usepackage{textcomp}
+\usepackage{multicol}
 \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
@@
 % List, can we do itemize for this? two columns? needs to be smaller
-Anthony Garrison, President
-Amanda Gaetano, Vice President
-Pat Hickey, Treasurer and Electronics Team Leader
-Greg Quinn, Mechanical Team Leader
-Stephen Indyk, Cogan Noll, John Palmer, Adrien Perkins, Michael Varga, Bradley Lord, David Wescott, and Taras Wallace.
+
+\begin{multicols}{2}
+\begin{itemize}
+	\setlength{\itemsep}{0cm}
+	\setlength{\parskip}{0cm}
+%	\centering
+	\item Anthony Garrison, President
+	\item Amanda Gaetano, Vice President
+	\item Pat Hickey, Treasurer and Electronics Team Leader
+	\item Greg Quinn, Mechanical Team Leader
+%\end{itemize}
+%\begin{multicols}{2}
+%\begin{itemize}
+	\setlength{\itemsep}{0cm}
+	\setlength{\parskip}{0cm}
+%	\centering
+	\item Stephen Indyk 
+	\item Cogan Noll
+	\item John Palmer 
+	\item Adrien Perkins
+	\item Michael Varga 
+	\item Bradley Lord
+	\item David Wescott
+	\item Taras Wallace
+\end{itemize}
+\end{multicols}
 
@@
 \subsection{Camera}
-
-MICHAEL : physical specification of camera (size, 5 megapixel stills). We selected it because...
-
-MICHAEL: camera interfaces: serial and USB
-
-The camera will be mounted in a pan and tilt unit, which will ... STEVE, PAT
-
-\subsection{Pan Tilt Unit}
 \begin{figure}
@@
 \end{figure}
+MICHAEL : physical specification of camera (size, 5 megapixel stills). We selected it because...
+
+MICHAEL: camera interfaces: serial and USB
+
+The camera will be mounted in a pan and tilt unit, which will ... STEVE, PAT
+
+\subsection{Pan Tilt Unit}
 
 \begin{figure}
-
+	\centering
+	\subfloat[Isometric View]{\label{fig:pantilt_isometric} 
+			\includegraphics[width=0.9\textwidth]{../images/pantilt_isometric.png}}
+\\
+	\subfloat[Side View]{\label{fig:pantilt_side} 
+			\includegraphics[width=0.9\textwidth]{../images/pantilt_side.png}}
+	\caption{Pan Tilt Unit}
+	\label{fig:pantilt}
+\end{figure}
+\begin{figure}
+	\subfloat[Linkage Detail View]{\label{fig:pantilt_linkdetail} 
+			\includegraphics[width=0.6\textwidth]{../images/pantilt_linkdetail.png}}
+	\subfloat[Linkage Velocity Plots]{\label{fig:pantilt_linkdetail} 
+			\includegraphics[width=0.4\textwidth]{../images/pantilt_linkplot.png}}
+	\caption{Tilt Linkage Detail}
+	\label{fig:pantilt_link}
 \end{figure}

### 2010-05-22 paper/paper.tex
added figure of knight two, finished 12ft telemaster

@@
 \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
-\hyphenation{auto-nomous}
+\hyphenation{auto-nomous construction}
 
@@
 
-Figure goes here: Photos of Knight Two in construction
-
+\begin{figure}
+	\centering
+	\subfloat[The unfinished \emph{Knight Two}]{\label{fig:knighttwo_construction} 
+			\includegraphics[width=0.4\textwidth]{../images/knighttwo_underconstruction.jpg}}
+	\subfloat[A completed 12 Foot Telemaster. Photo Credit: Hobby Lobby] {\label{fig:knighttwo_12ft_telemaster} 
+			\includegraphics[width=0.4\textwidth]{../images/telemaster_12ft_hobbylobby.jpg}}
+  	\label{fig:knighttwo}
+  	\caption[Knight Two]{Knight Two}
+\end{figure}
 Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations

### 2010-05-23 paper/paper.tex
added pre-flight procedures, more list fiddling

@@
 \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
+\newcommand{\itemspace}{	\setlength{\itemsep}{0cm} \setlength{\parskip}{0cm}}
 \hyphenation{auto-nomous construction}
@@
 
-\begin{multicols}{2}
-\begin{itemize}
-	\setlength{\itemsep}{0cm}
-	\setlength{\parskip}{0cm}
-%	\centering
-	\item Anthony Garrison, President
-	\item Amanda Gaetano, Vice President
-	\item Pat Hickey, Treasurer and Electronics Team Leader
-	\item Greg Quinn, Mechanical Team Leader
-%\end{itemize}
 %\begin{multicols}{2}
 %\begin{itemize}
-	\setlength{\itemsep}{0cm}
-	\setlength{\parskip}{0cm}
-%	\centering
-	\item Stephen Indyk 
-	\item Cogan Noll
-	\item John Palmer 
-	\item Adrien Perkins
-	\item Michael Varga 
-	\item Bradley Lord
-	\item David Wescott
-	\item Taras Wallace
-\end{itemize}
-\end{multicols}
+%	\setlength{\itemsep}{0cm}
+%	\setlength{\parskip}{0cm}
+%	\item Anthony Garrison, President
+%	\item Amanda Gaetano, Vice President
+%	\item Pat Hickey, Treasurer and Electronics Team Leader
+%	\item Greg Quinn, Mechanical Team Leader
+%	\item Stephen Indyk 
+%	\item Cogan Noll
+%	\item John Palmer 
+%	\item Adrien Perkins
+%	\item Michael Varga 
+%	\item Bradley Lord
+%	\item David Wescott
+%	\item Taras Wallace
+%\end{itemize}
+%\end{multicols}
+
+\vspace{12 pt}
+
+\begin{tabular}{l l}
+Anthony Garrison, President & Cogan Noll \\
+Amanda Gaetano, Vice President & John Palmer\\
+Pat Hickey, Treasurer and Electronics Team Leader & Adrien Perkins\\
+Greg Quinn, Mechanical Team leader & Michael Varga\\
+Stephen Indyk & (Bradley Lord OR Taras Wallace)\\
+\end{tabular}
+
+\vspace{24 pt}
 
@@
 
+\newpage
+\appendix
+\renewcommand\thesection{Appendix \Alph{section}}
+\section{Pre-flight Procedures}
+\begin{enumerate}
+	\setlength{\itemsep}{0cm}
+	\setlength{\parskip}{0cm}
+	\item Assemble plane.
+	\item Check all mechanical and electrical connections.
+		\begin{itemize}
+		\itemspace
+		\item Are all nuts/bolts loc-tited?
+		\item Are all servo extentions connected and properly secured?
+		\item Are all access hatches secured?
+		\end{itemize}
+	\item Position wingmen*, centerman**.
+	\item Pilot turns on transmitter, announces ``transmitter on.''
+	\item Pilot requests ``reciever on''.  Centerman turns on reciever, acknowledges.
+	\item Pilot completes surface checks:
+		\begin{itemize}
+		\itemspace
+		\item Ailerons
+		\item Elevator
+		\item Rudder
+		\item Throttle
+		\end{itemize}
+	\item Prime engine.
+	\item Pilot obtains visual and verbal confirmation that wingmen, centerman are in place and alert.
+	\item Pilot requests ``ignition on.''  Centerman moves ignition switch to the ``on'' position and acknowledges.
+	\item Starter requests ``prop clear.''  Starter obtains visual confirmation that no person is standing in the prop arc and announces ``prop clear.''
+	\item Starter announces ``contact'' and starts the engine.
+	\item Starter clears the area.
+	\item Pilot repeats surface and throttle checks at close range.
+	\item Pilot leaves starting area and performs surface and throttle checks from a distance (range test).
+	\item Pilot returns to starting area to prepare for taxi.
+	\item Pilot requests ``wingmen off.''  Wingmen carefully clear the area and acknowledge.
+	\item Pilot requests ``centerman off.''  Centerman carefully clears the area and acknowledge.
+	\item Pilot taxis airplane to appropriate location to begin takeoff.
+	\item Pilot commences takeoff.
+
+\end{enumerate}
+
 \bibliographystyle{plain}

### 2010-05-23 paper/paper.tex
footnotes acting funky, but look fine, so whatever for now
changed order of wingmen/centerman off, as per adrien

@@
 
-The team is made of 12 Rutgers undergraduates
+The team is made of 10 Rutgers undergraduates
 % List, can we do itemize for this? two columns? needs to be smaller
@@
 		\end{itemize}
-	\item Position wingmen*, centerman**.
+	\item Position wingmen\footnotemark, centerman\footnotemark.
 	\item Pilot turns on transmitter, announces ``transmitter on.''
@@
 		\end{itemize}
-	\item Prime engine.
+	\item Starter primes engine.
 	\item Pilot obtains visual and verbal confirmation that wingmen, centerman are in place and alert.
@@
 	\item Pilot returns to starting area to prepare for taxi.
+	\item Pilot requests ``centerman off.''  Centerman carefully clears the area and acknowledges.
 	\item Pilot requests ``wingmen off.''  Wingmen carefully clear the area and acknowledge.
-	\item Pilot requests ``centerman off.''  Centerman carefully clears the area and acknowledge.
 	\item Pilot taxis airplane to appropriate location to begin takeoff.
@@
 
+\footnotetext[1]{Two men, one positioned at each wing tip, preventing the plane from moving during engine start and subsequent surface checks.}
+\footnotetext{Man straddling the fuse, preventing the plane from moving during engine start and subsequent surface checks}
+
 \bibliographystyle{plain}

### 2010-05-23 paper/paper.tex
removed old team list (present one is good), added most of anthony's information to existing sections

@@
 
-The team is made of 10 Rutgers undergraduates
-% List, can we do itemize for this? two columns? needs to be smaller
-
-%\begin{multicols}{2}
-%\begin{itemize}
-%	\setlength{\itemsep}{0cm}
-%	\setlength{\parskip}{0cm}
-%	\item Anthony Garrison, President
-%	\item Amanda Gaetano, Vice President
-%	\item Pat Hickey, Treasurer and Electronics Team Leader
-%	\item Greg Quinn, Mechanical Team Leader
-%	\item Stephen Indyk 
-%	\item Cogan Noll
-%	\item John Palmer 
-%	\item Adrien Perkins
-%	\item Michael Varga 
-%	\item Bradley Lord
-%	\item David Wescott
-%	\item Taras Wallace
-%\end{itemize}
-%\end{multicols}
-
-\vspace{12 pt}
-
+The team is made of 10 Rutgers undergraduates:
 \begin{tabular}{l l}
@@
 Amanda Gaetano, Vice President & John Palmer\\
-Pat Hickey, Treasurer and Electronics Team Leader & Adrien Perkins\\
+Pat Hickey, Electronics Team Leader & Adrien Perkins\\
 Greg Quinn, Mechanical Team leader & Michael Varga\\
-Stephen Indyk & (Bradley Lord OR Taras Wallace)\\
+Stephen Indyk & Taras Wallace\\
 \end{tabular}
@@
  Larry Kosinar,
- Bill (last name),
+ Bill Perkowski,
  Mike Korb,
@@
 We have also nearly completed a very similar backup airframe which serve as a backup in case the \emph{Daedalus} has a disaster.
-The \emph{Knight Two} will be fitted with nearly identical R/C flight electronics, and accept the same autopilot and imaging system payload as the \emph{Daedalus}
+The \emph{Knight Two} will be fitted with nearly identical R/C flight electronics, and accept the same autopilot and imaging system payload as the \emph{Daedalus}.
 
@@
 We stripped and rebuilt nearly the entire airframe, adding carbon-fiber reinforcements to the wing in anticipation of increased wing loading due to our payload.
+Once rebuilt and covered, we tested the \emph{Daedalus} weighed 19.5 lbs with engine.
+A full 24oz fuel tank, plus 3.5 lbs electronics payload, would bring the maximum takeoff weight to 25.5lbs.
 
-Once rebuilt and covered, we tested the \emph{Daedalus} in our shop to ensure the rebuilt structure was strong enough.
-Of particular concern was interface between the wings and the center section---it was important to confirm that the wings were mounted and reinforced in a manner that would permit them to carry the load of the plane while in flight.
-To this end, we put the \emph{Daedalus} through a series of static and dynamic ``sandbag tests.''  The plane was carefully supported under the center section, nose, and tail while pre-weighted bags of sand were slowly placed symetrically and simultaneously on the underside of each wing, starting from the innermost rib.
-The \emph{Daedalus} easily supported XX lbs (XX percent of its weight) under static loading.
-We then lifted the plane off its supports and jostled it to simulate in-flight motion.
-The plane seemed to remain stable, and upon further observation after unloading, no cracks or defects were detected.
+We wanted to certify the wings and wing root would withstand large dynamic forces in flight. 
+We designed a test similar to mimic the 
+Federal Airworthiness Standards' structural requirments\footnote{Special Federal Aviation Regulation No.
+23, Subpart C--Structure.
+Assuming full load on wing as described in 23.335 paragraph 4.i., where an aircraft must pull out of a 7.5\textdegree dive at cruise velocity and cruise power ``with a load factor of 1.5 (0.5$g$.
+accelleration increment)'' \cite{far23}}.
+Our first test was designed to certify that the wing, supported at the wing root and fuselage, could take a bottom loading present in a 1.5$g$ accelleration, with a saftey factor of 1.5\footnote{As stated in 23.303 \cite{far23}.}.
 
-Satisfied, we proceeded through the usual engine and surface tests before taxiing the plane around a parking lot.
-We brought the plane back inside, examined it for signs of stress and, finding everything in good condition, deemed the \emph{Daedalus} ready to fly.
+The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symetrically and simultaneously on the underside of each wing, starting from the innermost rib.
+The \emph{Daedalus} wings supported 64 lbs (250\% of the maximum take-off weight) under static loading.
+We also rocked the plane ten degrees from side to side while fully loaded to simulate differential loads during flight.
+Our post-test structure inspection found no warped members or cracked joints.
+
+A second test was designed to certify the wing could take a top loading present from a 1.5$g$ accelleration with a saftey factor of 1.5, and to certify that load could be transmitted to the ground through the fuselage and landing gear.
+With the aircraft fully assembled and sitting upright on the landing gear, we loaded the top of the wing along the quarter chord with 64 lbs of sandbags.
+As in our first test, we also rocked the plane 10\textdegree  side to side to simulate uneven loading.
+Our post-test structure inspection found no warped members or cracked joints.
+
+Satisfied, made engine and surface tests (see \ref{sec:appendix_preflight}) before taxiing the plane around a parking lot.
+We brought the plane back inside, examined it for signs of stress and, finding everything in good condition, deemed the \emph{Daedalus} flightworthy.
 
@@
 
-Each airframe has a 45cc gasoline engine.
-Each airframe has a xx ounce gasoline tank.
-Based on our flight tests, with climbs to 400 feet AGL, a full xx ounce gasoline tank gives us xx minutes of flight time.
-(can someone guestimate this?
-Do we have data from break-in, etc)
+Each airframe has a 45cc gasoline engine and 24 ounce gasoline tank.
+Based on our flight tests, with climbs to 400 feet AGL and a high throttle cruise, we use a little less than 1oz gasoline per minute.
+While actual fuel consumption may be less, we set a flight time cap at 20 minutes in order to always have a saftey reserve of at least 4oz.
+We have tested to ensure that, in level flight, the fuel tank will feed the engine until less than $1/2$oz fuel remains.
 
@@
 \section{Pre-flight Procedures}
+\label{sec:appendix_preflight}
 \begin{enumerate}

### 2010-05-23 paper/paper.tex
made team list all 12 final members

@@
 
-The team is made of 10 Rutgers undergraduates:
-\begin{tabular}{l l}
-Anthony Garrison, President & Cogan Noll \\
-Amanda Gaetano, Vice President & John Palmer\\
-Pat Hickey, Electronics Team Leader & Adrien Perkins\\
-Greg Quinn, Mechanical Team leader & Michael Varga\\
+The team is made of 12 Rutgers undergraduates:\\
+\\ \begin{tabular}{l l}
+Anthony Garrison, President & John Palmer \\
+Amanda Gaetano, Vice President & Adrien Perkins\\
+Pat Hickey, Electronics Team Leader & Michael Varga\\
+Greg Quinn, Mechanical Team leader & Bradley Lord\\
 Stephen Indyk & Taras Wallace\\ 
+Cogan Noll & David Wescott\\ \\
 \end{tabular}
-
-\vspace{24 pt}
 

### 2010-05-24 paper/paper.tex
switched \textdegrees to new \degrees{} command. made steve's figs narrower so they fit on one page

@@
 \newcommand{\itemspace}{	\setlength{\itemsep}{0cm} \setlength{\parskip}{0cm}}
+\newcommand{\degrees}[1]
+{
+\begin{math}
+#1^{\circ} 
+\end{math}
+}
 \hyphenation{auto-nomous construction}
@@
 23, Subpart C--Structure.
-Assuming full load on wing as described in 23.335 paragraph 4.i., where an aircraft must pull out of a 7.5\textdegree dive at cruise velocity and cruise power ``with a load factor of 1.5 (0.5$g$.
+Assuming full load on wing as described in 23.335 paragraph 4.i., where an aircraft must pull out of a \degrees{7.5} dive at cruise velocity and cruise power ``with a load factor of 1.5 (0.5$g$.
 accelleration increment)'' \cite{far23}}.
@@
 With the aircraft fully assembled and sitting upright on the landing gear, we loaded the top of the wing along the quarter chord with 64 lbs of sandbags.
-As in our first test, we also rocked the plane 10\textdegree  side to side to simulate uneven loading.
+As in our first test, we also rocked the plane \degrees{10} side to side to simulate uneven loading.
 Our post-test structure inspection found no warped members or cracked joints.
@@
 While actual fuel consumption may be less, we set a flight time cap at 20 minutes in order to always have a saftey reserve of at least 4oz.
-We have tested to ensure that, in level flight, the fuel tank will feed the engine until less than $1/2$oz fuel remains.
+We have tested to ensure that, within \degrees{\pm10} of level flight, the fuel tank will feed the engine until less than $1/2$oz fuel remains.
 
@@
 	\subfloat[Isometric View]{\label{fig:pantilt_isometric} 
-			\includegraphics[width=0.9\textwidth]{../images/pantilt_isometric.png}}
+			\includegraphics[width=0.75\textwidth]{../images/pantilt_isometric.png}}
 \\
-	\subfloat[Side View]{\label{fig:pantilt_side} 
-			\includegraphics[width=0.9\textwidth]{../images/pantilt_side.png}}
+	\subfloat[Side View.]{\label{fig:pantilt_side} 
+			\includegraphics[width=0.75\textwidth]{../images/pantilt_side.png}}
 	\caption{Pan Tilt Unit}
@@
 We took advantage of the MHS-PM5's swiveling lens assembley 
-(see Figure \ref{fig:camera_frontview}, and added a servo actuator which could rotate the lens 135\textdegree back from straight forward, as shown in Figure \ref{}
+(see Figure \ref{fig:camera_frontview}, and added a servo actuator which could rotate the lens \degrees{135} back from straight forward, as shown in Figure \ref{fig:pantilt_linkdetail}.
+The tilt servo is a Hitech HS 77BB \ref{hitechs77bb}, and is capable of \degrees{180} rotation.
 The camera and tilt servo assembley is attached to the aircraft on a rotating mount.
+A thrust bearing takes lateral loads off the pan servo shaft, and stabilizes the assembley.
+The pan servo is a Hitech HS 785HB \ref{hitechhs785}, and is capable of \degrees{720} rotation.
+Both servos are powered and controlled by the Camera Control board.
+
+The total height of the imaging system is 8.75", only 4'' of which is exposed below the fuselage of the aircraft.
+The rotating assembly is, on its widest dimension 2.13", and would fit inside a 2.5" cylinder.
+Wires exiting the side of the camera, inside the fuselage, may make the assebley a bit wider.
+Because the maximum rotation is only \degrees{270}, the camera and tilt servo wires only have to be capable of wrapping \degrees{\pm 360} around the shaft below the thrust bearing.
+A wire guide will be added to keep the wires from tangling once construction of the pan-tilt is complete.
 

### 2010-05-24 paper/paper.tex
added battery details

@@
 Each airframe has nearly identical flight electronics.
-To power a Futaba 2.4GHz receiver and 6 servos for a XX minute flight, we selected a 2 cell XX amp hour Lithium Ion battery with a 5v, 10A capacity switching regulator made by Castle Creations.
-Pat will add more here if he has to.
+To power a Futaba 2.4GHz receiver and 6 servos for a 20 minute flight, we selected a 2 cell (7.2v) 3.2 amp hour Lithium Ion battery with a 5v, 10A capacity switching regulator made by Castle Creations.
+We use a low voltage monitor in-line with the battery, which beeps when cell voltage falls below 7.2v.
+The autopilot electronics are powered off a completely separate battery for safety.
 

### 2010-05-24 paper/paper.tex
spel chek

@@
 
-For this entry, we took a close look at our goals and requirments, and dialed back our approach to ensure we would meet the most basic competition requirments successfully.
+For this entry, we took a close look at our goals and requirements, and dialed back our approach to ensure we would meet the most basic competition requirements successfully.
 We started the design with some broad goals based on our experience in the 2009 competition. 
@@
 
-We elected to persue a kit-type airplane rather than our own design. 
+We elected to pursue a kit-type airplane rather than our own design. 
 Our own custom design, the \emph{Icarus}, suffered a catastrophic structural failure last year.
 From this, we learned that custom structures require much more testing and revision than a proven kit.
-We also reconsidered our payload requirments and decided we could easily modify a kit plane to accept our autopilot and imaging system.
+We also reconsidered our payload requirements and decided we could easily modify a kit plane to accept our autopilot and imaging system.
 
@@
 A member of our local R/C club, Tri-County RC \cite{tricountyRC}, donated a ten-foot wingspan high wing trainer to our club.
-The plane, first built out of foam and balsa from long-lost plans in 1986, was accepted graciously, but required quite a bit of work before it was once again flightworthy.
+The plane, first built out of foam and balsa from long-lost plans in 1986, was accepted graciously, but required quite a bit of work before it was once again flight-worthy.
 We removed the covering to find water damage and rot which resulted from years of storage.
@@
 We designed a test similar to mimic the 
-Federal Airworthiness Standards' structural requirments\footnote{Special Federal Aviation Regulation No.
+Federal Airworthiness Standards' structural requirements\footnote{Special Federal Aviation Regulation No.
 23, Subpart C--Structure.
 Assuming full load on wing as described in 23.335 paragraph 4.i., where an aircraft must pull out of a \degrees{7.5} dive at cruise velocity and cruise power ``with a load factor of 1.5 (0.5$g$.
-accelleration increment)'' \cite{far23}}.
-Our first test was designed to certify that the wing, supported at the wing root and fuselage, could take a bottom loading present in a 1.5$g$ accelleration, with a saftey factor of 1.5\footnote{As stated in 23.303 \cite{far23}.}.
+acceleration increment)'' \cite{far23}}.
+Our first test was designed to certify that the wing, supported at the wing root and fuselage, could take a bottom loading present in a 1.5$g$ acceleration, with a safety factor of 1.5\footnote{As stated in 23.303 \cite{far23}.}.
 
-The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symetrically and simultaneously on the underside of each wing, starting from the innermost rib.
+The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symmetrically and simultaneously on the underside of each wing, starting from the innermost rib.
 The \emph{Daedalus} wings supported 64 lbs (250\% of the maximum take-off weight) under static loading.
@@
 
-A second test was designed to certify the wing could take a top loading present from a 1.5$g$ accelleration with a saftey factor of 1.5, and to certify that load could be transmitted to the ground through the fuselage and landing gear.
+A second test was designed to certify the wing could take a top loading present from a 1.5$g$ acceleration with a safety factor of 1.5, and to certify that load could be transmitted to the ground through the fuselage and landing gear.
 With the aircraft fully assembled and sitting upright on the landing gear, we loaded the top of the wing along the quarter chord with 64 lbs of sandbags.
 As in our first test, we also rocked the plane \degrees{10} side to side to simulate uneven loading.
-Our post-test structure inspection found no warped members or cracked joints.
+Our post-test structure inspection once again found no warped members or cracked joints.
 
 Satisfied, made engine and surface tests (see \ref{sec:appendix_preflight}) before taxiing the plane around a parking lot.
-We brought the plane back inside, examined it for signs of stress and, finding everything in good condition, deemed the \emph{Daedalus} flightworthy.
+We brought the plane back inside, examined it for signs of stress and, finding everything in good condition, deemed the \emph{Daedalus} flight-worthy.
 
@@
 Based on our flight tests, with climbs to 400 feet AGL and a high throttle cruise, we use a little less than 1oz gasoline per minute.
-While actual fuel consumption may be less, we set a flight time cap at 20 minutes in order to always have a saftey reserve of at least 4oz.
+While actual fuel consumption may be less, we set a flight time cap at 20 minutes in order to always have a safety reserve of at least 4oz.
 We have tested to ensure that, within \degrees{\pm10} of level flight, the fuel tank will feed the engine until less than $1/2$oz fuel remains.
@@
 
-\subsection{Requirments}
+\subsection{Requirements}
 
@@
 
-We took advantage of the MHS-PM5's swiveling lens assembley 
+We took advantage of the MHS-PM5's swiveling lens assembly 
 (see Figure \ref{fig:camera_frontview}, and added a servo actuator which could rotate the lens \degrees{135} back from straight forward, as shown in Figure \ref{fig:pantilt_linkdetail}.
@@
 The camera and tilt servo assembley is attached to the aircraft on a rotating mount.
-A thrust bearing takes lateral loads off the pan servo shaft, and stabilizes the assembley.
+A thrust bearing takes lateral loads off the pan servo shaft, and stabilizes the assembly.
 The pan servo is a Hitech HS 785HB \ref{hitechhs785}, and is capable of \degrees{720} rotation.
@@
 The rotating assembly is, on its widest dimension 2.13", and would fit inside a 2.5" cylinder.
-Wires exiting the side of the camera, inside the fuselage, may make the assebley a bit wider.
+Wires exiting the side of the camera, inside the fuselage, may make the assembly a bit wider.
 Because the maximum rotation is only \degrees{270}, the camera and tilt servo wires only have to be capable of wrapping \degrees{\pm 360} around the shaft below the thrust bearing.
@@
 As such, we've implemented numerous safety features and developed a strict procedure for starting, taxiing, and flying the plane.
-This procedure enables us to ensure that all personnel are safely in place, and that no intergral checks are missed.
+This procedure enables us to ensure that all personnel are safely in place, and that no integral checks are missed.
 
@@
 	\item Channel 5 (?) allows quick switch to safety pilot in case of autopilot malfunction.
-	\item If the plane loses contact with the ground station for more than XX seconds, the autopilot will default to a ``return home mode'' that will direct the plane towards the groundstation in hopes of reestablishing a connection.
+	\item If the plane loses contact with the ground station for more than XX seconds, the autopilot will default to a ``return home mode'' that will direct the plane towards the ground station in hopes of reestablishing a connection.
 If the connection remains dead for XX minutes, the autopilot will direct the plane to [DO SOMETHING?]
@@
 		\item Are all nuts/bolts loc-tited?
-		\item Are all servo extentions connected and properly secured?
+		\item Are all servo extensions connected and properly secured?
 		\item Are all access hatches secured?
@@
 	\item Pilot turns on transmitter, announces ``transmitter on.''
-	\item Pilot requests ``reciever on''.  Centerman turns on reciever, acknowledges.
+	\item Pilot requests ``receiver on''.  Centerman turns on receiver, acknowledges.
 	\item Pilot completes surface checks:

### 2010-05-24 paper/paper.tex
eliminated abstract (page limit). finished up pantilt writing

@@
 \hypersetup{ colorlinks=true
-           , linkcolor=cyan
+           , linkcolor=blue
            , pdfnewwindow=true 
@@
 
-\begin{abstract}
-We will write the abstract after the rest of the paper is finished. If we're really close to the page limit, we may be able to omit it.
-\end{abstract}
+% We are definitely going to hit the page limit. No more abstract.
+%\begin{abstract}
+%We will write the abstract after the rest of the paper is finished. 
+%If we're really close to the page limit, we may be able to omit it.
+%\end{abstract}
 
@@
 			\includegraphics[width=0.6\textwidth]{../images/pantilt_linkdetail.png}}
-	\subfloat[Linkage Velocity Plots]{\label{fig:pantilt_linkdetail} 
+	\subfloat[Linkage Velocity Plots]{\label{fig:pantilt_linkplots} 
 			\includegraphics[width=0.4\textwidth]{../images/pantilt_linkplot.png}}
@@
 The tilt servo is a Hitech HS 77BB \ref{hitechs77bb}, and is capable of \degrees{180} rotation.
+Figures \ref{fig:pantilt} and \ref{fig:pantilt_linkdetail} show the camera lens tilted at \degrees{45} below the horizon.
 The camera and tilt servo assembley is attached to the aircraft on a rotating mount.
 A thrust bearing takes lateral loads off the pan servo shaft, and stabilizes the assembly.
-The pan servo is a Hitech HS 785HB \ref{hitechhs785}, and is capable of \degrees{720} rotation.
+The pan servo is a Hitech HS 785HB \ref{hitechhs785}, and is capable of \degrees{1260} rotation (3.5 revolutions).
 Both servos are powered and controlled by the Camera Control board.
@@
 Wires exiting the side of the camera, inside the fuselage, may make the assembly a bit wider.
-Because the maximum rotation is only \degrees{270}, the camera and tilt servo wires only have to be capable of wrapping \degrees{\pm 360} around the shaft below the thrust bearing.
-A wire guide will be added to keep the wires from tangling once construction of the pan-tilt is complete.
+Because the maximum rotation is only \degrees{1260}, the camera and tilt servo wires only have to be capable of wrapping \degrees{\pm 630} around the shaft below the thrust bearing.
+A wire guide will be added to keep the wires from tangling once construction of the pan-tilt is complete and installed in the aircraft.
+
+The pan servo is calibrated before installation by mapping servo PWM inputs to a measured angular output at \degrees{30} intervals.
+Linear interpolation is used to determine PWM outputs which achieve intermediate angles.
+The tilt servo is calibrated at \degrees{10} intervals, due to its much smaller rotation range.
+As shown in Figure \ref{fig:pantilt_linkplots}, the angular accelleration of the camera angle linkage (Alpha4) is almost constant compared to the servo arm linkage throughout the range $[0^{\circ}, 135^{\circ}]$.
+The linkage has been designed to keep the majority of nonlinear rotation outside the camera's range of rotation.
+Therefore, we can consider the calibrated servo output to be directly related to the camera angle.
+The endpoints of the tilt rotation rang are calibrated and enforced in software, because pushing the camera outside this range may damage it.
 
@@
 \renewcommand\thesection{Appendix \Alph{section}}
-\section{Pre-flight Procedures}
+\section{Preflight Procedures}
 \label{sec:appendix_preflight}

### 2010-05-24 paper/paper.tex
added some revisions of michael's work. added some comments to cogan's work.

@@
 \end{figure}
-MICHAEL : physical specification of camera (size, 5 megapixel stills). We selected it because...
+
+We selected the Sony MHS-PM5 camera as the basis of our imaging system.
+Marketed as a pocket 1080i video camera capable of 5 megapixel (MP) still images, we primarily selected the camera because it features a swiveling lens at one end of the body and it features Sony's LANC (Local Application Control Bus System, ref{lanc}).
+Because the swiveling lens is at one end of the body, we can mount the camera primarily inside the fuselage of the aircraft, which will protect it from damage on takeoff and landing.
+Internal mounting also helps reduce drag.
+The swiveling lens is an important feature because it allows us to achieve two rotational degrees of freedom while only rotating the body about one axis.
+This simplifies the pan tilt unit design.
+
+The LANC protocol is based on a 9600 baud UART serial connection, and is available on the PM5's 10-pin ``Multi-AV'' connector.
+Camera control commands, which include commands to zoom in, zoom out, and take a still picture, is publicly available \ref{boehmenl_lanc}.
+Also on the ``Multi-AV'' connector are the signals for NTSC video and audio output.
+We use both the NTSC output and LANC signals on this connector using a commercially available adapter cable \ref{lanc_multiav_cable}.
+
+% first citation should go to http://www.boehmel.de/lanc.htm
+
+We use the PM5 to transmit NTSC video to the ground, where it is viewed by an operator. The operator can wirelessly trigger a still picture, which is stored on the PM5's 2GB MemoryStickDuo card.
+
+\subsection{Camera-Computer Interface}
+
+ We can access the contents of that card via the PM5's USB connector.
+When attached to a computer, the PM5 powers the camera off and appears as a USB Mass Storage Device.
+In order to use the PM5 as both a camera and Mass Storage Device in flight, we have spliced a transistor into the power line on the USB cable.
+We control whether the camera detects it is connected via USB by blocking or allowing current in that line.
+
+%Overview of Ground Station Commands
+%Some of the commands that the ground station can issue to the plane are capture image, zoom in, zoom out, crop image, download crop, generate crop, pan left, pan right, tilt up and tilt down.
+These commands are initiate by a simple button press on the ground station gui program.
+The ground station sends a signal to the plane computer through a radio.
+The plane software then handles the command directly or may forward it on to the microcontroller.
+The capture image, pan and tilt commands are forwarded onto the microcontroller as they involve the lanc interface with the camera and the servos.
+The download crop and generate crop commands are handled directly by the computer since the data necessary to carry out these functions exists on its hard drive..
 
@@
 
-Once images are on the plane's computer they are ready to be downloaded over the Xbee Radio link.
-Each image is fairly large, and our bandwidth is only around 115200 baud, so we use a bandwidth conserving method to transfer images.
+Once images are on the plane's computer they are ready to be downloaded over a 2.4GHz XBee radio link.
+Each image is fairly large, and our bandwidth is only around 115200 baud % actually throughput, at worst-case, is half this.
+, so we select only the most important part of the images to download.
 The entire image is first downloaded as a low-res `thumbnail.'  Once that is available for viewing, the operator can select an area of interest and download a high-res crop of just that area.
 This allows us to quickly obtain high resolution images of the targets of interest.
+
+% Please state the sizes of the thumbnail and cropped images, compared to the total image size, and state the average download times of each, in a table. this will make our improvements dramatic.
 

### 2010-05-24 paper/paper.tex
finished all the rewriting of michael's camera description i want to for tonight. inserted his raw stuff as comment for cogan to review

@@
 
-We use the PM5 to transmit NTSC video to the ground, where it is viewed by an operator. The operator can wirelessly trigger a still picture, which is stored on the PM5's 2GB MemoryStickDuo card.
+The PM5's NTSC video signal outputs a 480 line video of the current ``viewfinder'', the live video that is also displayed on the rear LCD screen. 
+Still pictures are taken based on targets spotted in this video feed. Those images are stored on the camera's memory card.
 
@@
 
- We can access the contents of that card via the PM5's USB connector.
+We access the still pictures on the memory card via the PM5's USB connector.
 When attached to a computer, the PM5 powers the camera off and appears as a USB Mass Storage Device.
@@
 
+% cogan, please finish up this section. Some of the information may be redundant.
+% we also need a figure of this system, labeling all the components and connections.
+
+% from michael varga:
 %Overview of Ground Station Commands
@@
 The download crop and generate crop commands are handled directly by the computer since the data necessary to carry out these functions exists on its hard drive..
-
-MICHAEL: camera interfaces: serial and USB
-
-The camera will be mounted in a pan and tilt unit, which will ... STEVE, PAT
+%Target Verification
+%A live video feed is forwarded from the camera through a ntsc video transmitter.
+The ground station laptop contains a ntsc video receiver so it may be displayed on the screen.
+When the operator sees a target in the view of the camera he issues a command to the plane to capture an image of the camera’s current view.
+These images will be stored on the plane and ready for download when instructed by the ground station operator.
+%The original full resolution images are too large to be downloaded directly.
+A lower resolution picture is generated for every high resolution image that is captured.
+The ground station operator must download and look at the thumbnail to select a rectangular region that encompasses the target.
+A command is issued to the plane that generates a crop from the original higher resolution image using the selected rectangular region.
+The ground station operator may then download this newly generated crop to verify the target with more precision.
 
@@
 They view a live video stream coming from the camera on the plane, which is transmitted over a high bandwidth radio.
-The operator can also control the camera by sending commands to the plane over an Xbee Radio link.
+The operator can also control the camera by sending commands to the plane over the XBee radio link.
 In order to assist the operator in finding targets, a target locating algorithm will be run on the video feed to highlight possible targets.
@@
 When a picture is taken on the plane it is immediately saved on the camera's memory card.
-In order to view that picture on the ground it must first be transferred to the plane's main computer (beagle board), and then downloaded to the Image Station over the Xbee connection.
+In order to view that picture on the ground it must first be transferred to the plane's main computer (beagle board), and then downloaded to the Image Station over the XBee connection.
 

### 2010-05-24 paper/paper.tex
added comment on what else i need in camera section
reorganized systems engineering a bit, wrote airframe system engineering paragraph i want cogan to improve on

@@
 
-One of the most difficult design decisions was to quantify ``ample payload.'' The payload size and weight is the major determining factor in the aircraft size.
-% -- expand this thought pch: 
-We started off assuming an electronics (computer + batteries) payload of 2.5lbs and a pan-tilt camera of 2.5lbs and 8 inch diameter.
-
-% Originally a large and heavy system was to be used for an imaging system, but as the year progress a lighter and smaller alternative was found, which helped the mechanical team in the placement of the imaging system(I do not know the details of what they did, that’s all Pat)
+\subsubsection{Airframe System Engineering}
 
@@
 
-% -- donno how much use this is:
-%  During both constructions we split into a wing and tail section group and a fuselage group.
-While all the construction was going on, two expert builders from the local R/C club would come and advise us when we were unsure how to proceed in order to have the best balance of weight and strength.
+One of the most difficult design decisions was to quantify ``ample payload.'' The payload size and weight is the major determining factor in the aircraft size.
+% -- amanda will finish this by mon noon: 
+We started off assuming an electronics (computer + batteries) payload of 2.5lbs and a pan-tilt camera of 2.5lbs and 8 inch diameter.
 
-%Along with the actual construction of the planes, the mechanical team also focused on modifying the structures of the planes in order to fit the desired payload: the autopilot system and the imaging system.
+\subsubsection{Autopilot System Engineering}
 
 The electronics team was in charge of the autopilot system and the imaging system.
-% not sure what i want to do to expand here pch
+
+\subsubsection{Payload System Engineering}
+% Cogan, please explain the overall design decisions here. These are one sentence points in not great english, maybe make these two sentences each. 
+We used an off the shelf consumer camera because small, cheap, capable .
+We needed a linux based computer to interface with camera - only viable option, if we don't use a prebuilt system.
+We used our own pantilt unit because we wanted it to custom fit our camera.
+and it turned out our camera choice heavily effected our pantilt design, and allowed us to have a very compact system.
+ We used an independent radio link because we weren't sure about integration with the autopilot system, and we wanted to be able to switch the way the autopilot worked without dependencies.
+We wrote our own ground station software out of necessity, turned out to be excellent design decision.
 
@@
 Still pictures are taken based on targets spotted in this video feed. Those images are stored on the camera's memory card.
+
+% please give details of still pictures, including the field of view per pixel.
+include a quick table of resolving power (projection of pixel field of view, center of frame, onto ground) at altitudes [100,200,300,400 agl] for straight down and 45 degree angle.
+Please also give the resolving power of the video viewfinder
 

### 2010-05-24 paper/paper.tex
really expanding systems engineering

@@
 
+\subsection{Goals}
 For this entry, we took a close look at our goals and requirements, and dialed back our approach to ensure we would meet the most basic competition requirements successfully.
-We started the design with some broad goals based on our experience in the 2009 competition. 
-Sub-teams and individuals tackled the smaller problems while still keeping an eye on the big picture.
+We started the design by forming, as a team, some broad goals based on our experience in the 2009 competition.
+These formed the basis on which sub-teams and individuals tackled the smaller problems while still keeping an eye on the big picture.
 
-\subsubsection{Goals}
+We also wanted to balance building with buying, keeping in mind our slim finances.
+While its impossible to truly start from scratch\footnote{``If you want to make an apple pie from scratch, you must first create the universe'' - Dr.
+Carl Sagan \cite{saganquote}}, we tried to build whatever we could given the 16th June competition deadline.
+
+% footnote is broken (breaks from bottom of page one onto bottom of page two), something about the documentclass needs to be changed so it doesn't assume pages are next to each other (book-like)?
+
 \begin{itemize}
-	\item The air vehicle must be stable and slow, with ample payload capacity and 30 minute flight duration. It should be a proven and durable fixed-wing design.
-	\item The autopilot must be based on proven software, so we could focus on improving an existing system rather than reinventing the wheel.
-We felt the Paparazzi system was full featured and proven, and we wanted to adapt it to our needs.
+	\item The air vehicle must be stable and slow, with ample payload capacity and 20 minute flight duration. It should be a proven and durable fixed-wing design.
+	\item The autopilot must be based on a proven system, so we could focus on improving an existing system rather than reinventing the wheel.
+It also must be open-source, both because of the high cost commercial systems, and because we want to learn about and experiment with the internals of our autopilot system.
 	\item The imaging system must be simple and capable.
 We assume the operator will be burdened with guiding the camera and identifying targets until time and resources permit us to add sensor information and control loops for guidance, and computer vision for identification.
+It will be custom built, because we could not find any affordable systems available commercially.
 \end{itemize}
@@
 We decided not to pursue the reach goals of autonomous takeoff and landing based on our inexperience and the risk, during development and competition, to our aircraft.
-
+\subsection{System Design Decisions}
 \subsubsection{Airframe System Engineering}
@@
 
+\subsubsection{Electronics System Engineering}
+
+The electronics team was in charge of the autopilot system and the imaging system.
+The biggest system design choice we made was to use a single board Linux-based flight computer as the platform for both systems.
+We felt that the best way to develop software which interfaces with a wide variety of analog and digital electrical systems, has complicated information flow requirements,
+
+We felt the Paparazzi system was full featured and proven, and we wanted to adapt it to our needs.
+
+\subsubsection{Safety of Linux Based Flight Computer}
+
+We acknowledge that a Linux based flight computer does not support the requirements for hard real-time, deterministic computing that would be required of any real-world autonomous system.
+Given the resources available to the team (time, money, and manpower), such a fully assured system is out of reach.
+
+We elected to take many steps in the software engineering to help ensure our flight computer will operate reliably.
+The most important of these was code review and incremental testing.
+Both Pat Hickey and Bradley Lord became intimately familiar with the source code and control flow inside the autopilot software, and all patches were reviewed and tested by both engineers.
+Numerous other safety measures will be described in Section \ref{sec:autopilot_software}.
+
+Because our entry is essentially an experimental system, and will always be operated with a safety pilot present, we engineered the safety pilot override to operate independently from the flight computer.
+We have ensured that, short of losing power to the radio control receiver and servos, the safety pilot will always be able to take over at a moment's notice.
+We have also ensured that the flight computer's servo controller will be able to detect a loss of computer control and servos to a recoverable default.
+Along with careful preflight testing, careful observation, and definitive analysis of any errors encountered during flight, these measures meet the spirit and letter of competition rules.
+
 \subsubsection{Autopilot System Engineering}
 
-The electronics team was in charge of the autopilot system and the imaging system.
+As stated previously, the Paparazzi system is the most capable and proven open-source autopilot system available.
+The first task was to select hardware which would allow us to port the airborne code from the Paparazzi's standard microcontroller-based autopilot hardware to our flight computer.
 
@@
 \subsection{Software}
+\label{sec:autopilot_software}
 PAT
@@
 The tilt servo is calibrated at \degrees{10} intervals, due to its much smaller rotation range.
-As shown in Figure \ref{fig:pantilt_linkplots}, the angular accelleration of the camera angle linkage (Alpha4) is almost constant compared to the servo arm linkage throughout the range $[0^{\circ}, 135^{\circ}]$.
+As shown in Figure \ref{fig:pantilt_linkplots}, the angular acceleration of the camera angle linkage (Alpha4) is almost constant compared to the servo arm linkage throughout the range $[0^{\circ}, 135^{\circ}]$.
 The linkage has been designed to keep the majority of nonlinear rotation outside the camera's range of rotation.

### 2010-05-24 paper/paper.tex
commented out sagan footnote for now. preflight footnotes in-line.

@@
 We also wanted to balance building with buying, keeping in mind our slim finances.
-While its impossible to truly start from scratch\footnote{``If you want to make an apple pie from scratch, you must first create the universe'' - Dr.
-Carl Sagan \cite{saganquote}}, we tried to build whatever we could given the 16th June competition deadline.
+While we couldn't build every part of our entry from scratch, we tried to build whatever we could given the 16th June competition deadline.
 
-% footnote is broken (breaks from bottom of page one onto bottom of page two), something about the documentclass needs to be changed so it doesn't assume pages are next to each other (book-like)?
+%\footnote{``If you want to make an apple pie from scratch, you must first create the universe'' - Dr. Carl Sagan \cite{saganquote}},
 
@@
 
+\subsection{Safety Observers}
+A safety observer is defined as a person, devoid of any additional specific responsibilities, who remains in the starting area during all pre-flight procedures.
+The safety observer is expected to watch for unsafe conditions, included but not limited to non-essential personnel in the starting, takeoff, and/or flight areas; loose mechanical connections; failed surface and/or range checks; and other forms of mechanical failure.
+The safety observer has the absolute authority to give the order to kill the ignition and abort startup immediately upon observing any unsafe condition.
+The safety observer is responsible for communicating possible conditions to the startup team as they arise.
+At minimum, one safety observer is required to start the engine, though we typically have more.
+
 \newpage
@@
 		\end{itemize}
-	\item Position wingmen\footnotemark, centerman\footnotemark.
+	\item Position wingmen\footnote{Two men, one positioned at each wing tip, preventing the plane from moving during engine start and subsequent surface checks.}, centerman\footnote{Man straddling the fuse, preventing the plane from moving during engine start and subsequent surface checks}.
 	\item Pilot turns on transmitter, announces ``transmitter on.''
@@
 
-\footnotetext[1]{Two men, one positioned at each wing tip, preventing the plane from moving during engine start and subsequent surface checks.}
-\footnotetext{Man straddling the fuse, preventing the plane from moving during engine start and subsequent surface checks}
-
 \bibliographystyle{plain}

### 2010-05-24 paper/paper.tex
amanda's latest airplane systems engineering

@@
 
-Our team was divided by area of expertise into a mechanical team and an electrical team.
-The mechanical team began the year with the goal of producing two competition worthy airframes. These were to become the \emph{Daedalus} and \emph{Knight Two}.
+Our team was divided by area of expertise into a mechanical team and an electrical team.The mechanical team began the year with the goal of producing two competition worthy airframes.
+These were to become the \emph{Daedalus} and \emph{Knight Two}.
+
+The teams made design decisions based on the aforementioned competition goals.
+
+One of the most difficult design decisions was to quantify ``ample payload.'' Payload size and weight is the major determining factor in aircraft size.
+We originally assumed a 2.5lb electronics (computer + batteries) payload and a 2.5lb pan-tilt camera 8'' in diameter, requiring 6'' of vertical space extruding from the bottom of the aircraft.
+These considerations introduced two major design parameters: the aircraft must be capable of carrying a minimum 5lb payload, and must be situated to provide clearance for the pan tilt unit.
+
+Additionally, we sought an easily controllable, stable platform to accommodate the autopilot, and an ability to fly at low speeds, in order to more easily capture clear images.
 
 We determined, after consulting with R/C experts, that a high wing trainer is the most stable and proven fixed-wing design.
-Trainers traditionally have light wing loading for slow speed flight, and will be able to carry payload
+Trainers traditionally have light wing loading for slow speed flight, and are able to carry a comparatively large payload.
+Additionally, a large plane with lower wing loading is easier to control, and easier to recover if power is lost.
+Consciously agreeing to err on the side of caution, we sought a large plane capable of accommodating a worst-case scenario payload, and settled on a wingspan on the order of 10'-12'.
+Though lighter and smaller payload alternatives were found as the year progressed, these merely leave us with more freedom for future modifications.
+%That last sentence sounds kind of eh, I'm stuck on a decent wording, so I'm leaving it for now.
 
-One of the most difficult design decisions was to quantify ``ample payload.'' The payload size and weight is the major determining factor in the aircraft size.
-% -- amanda will finish this by mon noon: 
-We started off assuming an electronics (computer + batteries) payload of 2.5lbs and a pan-tilt camera of 2.5lbs and 8 inch diameter.
+The decision to employ a large, high wing trainer, however, forced us to rethink our original pan-tilt design.
+Last year, in order to accommodate the pan-tilt unit, we designed a pusher plane, enabling us to mount the pan-tilt under the plane's nose, which had sufficient clearance.
+Since trainers are generally designed with front-mounted engines (I presume?), we were forced to engineer a new solution requiring significantly less clearance.
+In the long run, we modified the pan-tilt mount to allow us to recess more of the pan-tilt unit into the body of the plane---this had the added benefit of providing more protection to an expensive piece of equipment on takeoff and landing.
 

### 2010-05-24 paper/paper.tex
added wreckage figure, made intro page break, added louis, gregg peters to acknowledgments, dropped chris from advisors because we dont know his last name

@@
  Bill Perkowski,
- Mike Korb,
- (the man who donated daedalus), and
- Chris (test pilot) amongst many who spent time mentoring us.
+and  Mike Korb, 
+%and Chris (test pilot) 
+amongst many who spent time mentoring us.
 
@@
 We would like to acknowledge
+Louis Stumpf, for donating the aircraft which became the \emph{Daedalus};
 Jeff Perkins, for his prolific networking and promotion of the team;
-Gregg (last name) of BP Hobbies for getting us out of numerous binds;
+Gregg Peters of BP Hobbies for getting us out of numerous binds;
 Dave Devries and Michael Maia, of Invensense, for their technical guidance;
@@
 
+\begin{figure}[b]
+	\centering
+	\includegraphics[width=0.6\textwidth]{../images/wreckage.jpg}
+	\caption{The recovery of our 2009 entry, the \emph{Icarus}}
+	\label{fig:wreckage}
+\end{figure}
+
+\newpage
 \section{Systems Engineering}
-
 \subsection{Goals}

### 2010-05-24 paper/title.tex
added BP Hobbies
updated authors list based on final team list
switched to image of both planes, respaced so the image could be bigger

@@
 
-\includegraphics[width=0.5\textwidth]{../images/daedalus.jpg}\\[1cm]
-\HRule \\[1cm]
-{ \huge \bfseries Rutgers University Autonomous Aircraft Team } \\[0.5cm]
-\HRule \\[0.5cm]
+\includegraphics[width=0.6\textwidth]{../images/bothplanes_isometric.jpg}\\[7mm]
+\HRule \\[7mm]
+{ \huge \bfseries Rutgers University Autonomous Aircraft Team } \\[0.4cm]
+\HRule \\[0.4cm]
 { \large \bfseries Technical Report }
-\\[0.5cm]
+\\[0.4cm]
 { \large \bfseries 2010 AUVSI UAS Competition }
-\\[0.5cm]
-\HRule \\[1cm]
+\\[0.3cm]
+\HRule \\[7mm]
 
   {\large Amanda Gaetano, Anthony Garrison, Pat Hickey,}
-\\{\large Stephen Indyk, Cogan Noll, John Palmer,}
-\\{\large Adrien Perkins, Gregory Quinn, and Michael Varga}
+\\{\large Stephen Indyk, Bradley Lord, Cogan Noll,}
+\\{\large John Palmer, Adrien Perkins, Gregory Quinn,}
+\\{\large Michael Varga, Taras Wallace, and David Wescott}
 \\[1cm]
-\emph{Thank you to our sponsors}
+\emph{Thank you to our sponsors:}
 \\ Rutgers University Engineering Governance Council
@@
 \\ ST Micro, Inc.
+\\ BP Hobbies
 

### 2010-05-24 paper/paper.tex
added sandbag testing figure as a wrapfig.
working with figures, making some of them wrap

@@
-\documentclass[10pt]{report}
+\documentclass[10pt,twocolumns]{report}
 \usepackage[pdftex]{graphicx}
+\usepackage[left=1.5in, right=1.5in, top=1.25in, bottom=1.25in]{geometry}
 \usepackage{hyperref}
@@
 \usepackage{subfig}
+\usepackage{wrapfig}
 \usepackage{listings}
@@
 
-\begin{figure}[b]
+\begin{figure}[b!]
 	\centering
@@
 \subsubsection{Daedalus}
-
+\begin{wrapfigure}{l}{0.4\textwidth}
+	\includegraphics[width=0.37\textwidth]{../images/daedalus_isometric.jpg}
+	\caption{The \emph{Daedalus}}
+	\label{fig:daedalus}
+\end{wrapfigure}
 A member of our local R/C club, Tri-County RC \cite{tricountyRC}, donated a ten-foot wingspan high wing trainer to our club.
@@
 Our post-test structure inspection found no warped members or cracked joints.
-
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\includegraphics[width=0.37\textwidth]{../images/daedalus_sandbagtest.jpg}
+	\caption{Load testing the underside of the \emph{Daedalus}'s wings. Shown with 16lbs of load.}
+	\label{fig:sandbagtest}
+\end{wrapfigure}
 A second test was designed to certify the wing could take a top loading present from a 1.5$g$ acceleration with a safety factor of 1.5, and to certify that load could be transmitted to the ground through the fuselage and landing gear.
@@
 
-Figure goes here: Photos of Daedalus in construction, sandbag testing
+\begin{wrapfigure}{l}{0.4\textwidth}
+	\centering
+	\subfloat[The unfinished \emph{Knight Two}]{\label{fig:knighttwo_construction} 
+			\includegraphics[width=0.37\textwidth]{../images/knighttwo_uncovered_outdoors.jpg}}
+\\
+	\subfloat[A completed 12 Foot Telemaster. Photo Credit: Hobby Lobby] {\label{fig:knighttwo_12ft_telemaster} 
+			\includegraphics[width=0.37\textwidth]{../images/telemaster_12ft_hobbylobby.jpg}}
+  	\label{fig:knighttwo}
+  	\caption[Knight Two]{Knight Two}
 
+\end{wrapfigure}
 Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations
-
-Figure may go here: Line drawing of Daedalus. JOHN
 
@@
 
-\begin{figure}
-	\centering
-	\subfloat[The unfinished \emph{Knight Two}]{\label{fig:knighttwo_construction} 
-			\includegraphics[width=0.4\textwidth]{../images/knighttwo_underconstruction.jpg}}
-	\subfloat[A completed 12 Foot Telemaster. Photo Credit: Hobby Lobby] {\label{fig:knighttwo_12ft_telemaster} 
-			\includegraphics[width=0.4\textwidth]{../images/telemaster_12ft_hobbylobby.jpg}}
-  	\label{fig:knighttwo}
-  	\caption[Knight Two]{Knight Two}
-\end{figure}
 Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations

### 2010-05-25 paper/paper.tex
working on electronics system eignieering. moved a figure or something before. changed name of camera control board to pantilt control board.

@@
 
-We felt the Paparazzi system was full featured and proven, and we wanted to adapt it to our needs.
+We felt the Paparazzi system was full featured and proven, but we weren't satisified with the capabilities of the existing Paparazzi hardware. The ``Tiny'' v2.1\cite{paparazzi_tinyv21}, the latest
 
@@
 
-The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symmetrically and simultaneously on the underside of each wing, starting from the innermost rib.
-The \emph{Daedalus} wings supported 64 lbs (250\% of the maximum take-off weight) under static loading.
-We also rocked the plane ten degrees from side to side while fully loaded to simulate differential loads during flight.
-Our post-test structure inspection found no warped members or cracked joints.
  \begin{wrapfigure}{r}{0.4\textwidth}
@@
 \end{wrapfigure}
+The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symmetrically and simultaneously on the underside of each wing, starting from the innermost rib.
+The \emph{Daedalus} wings supported 64 lbs (250\% of the maximum take-off weight) under static loading.
+We also rocked the plane ten degrees from side to side while fully loaded to simulate differential loads during flight.
+Our post-test structure inspection found no warped members or cracked joints.
+
 A second test was designed to certify the wing could take a top loading present from a 1.5$g$ acceleration with a safety factor of 1.5, and to certify that load could be transmitted to the ground through the fuselage and landing gear.
@@
 The pan servo is a Hitech HS 785HB \ref{hitechhs785}, and is capable of \degrees{1260} rotation (3.5 revolutions).
-Both servos are powered and controlled by the Camera Control board.
+Both servos are powered and controlled by the Pan-tilt control board.
 

### 2010-05-25 paper/paper.tex

@@
 
-\subsubsection{Payload System Engineering}
-% Cogan, please explain the overall design decisions here. These are one sentence points in not great english, maybe make these two sentences each. 
-We used an off the shelf consumer camera because small, cheap, capable .
-We needed a linux based computer to interface with camera - only viable option, if we don't use a prebuilt system.
-We used our own pantilt unit because we wanted it to custom fit our camera.
-and it turned out our camera choice heavily effected our pantilt design, and allowed us to have a very compact system.
- We used an independent radio link because we weren't sure about integration with the autopilot system, and we wanted to be able to switch the way the autopilot worked without dependencies.
-We wrote our own ground station software out of necessity, turned out to be excellent design decision.
+\subsubsection{Imaging System Engineering}
+
+When designing the imaging system our major goals were to have 1) a camera with XXX range of motion.
+2) real-time video feed from the plane.
+3) ability to control the camera in real-time and take pictures.
+4) ability to download pictures while in flight.
+For the camera we chose the Sony Bloggie \cite{bloggie} because it was small, light, cheap, and took high resolution images and video.
+In addition it came with a built in rotating lens, which simplified the construcion of the pan-tilt unit, and it can also be controlled remotely over usb, so it interfaced with our flight linux computer.
+To communicate with the plane we developed custom ground station software to meet our needs and installed two sets of radios, one for the video feed, and another for communicating with the plane and downloading images.
+These radios are independent from the set of radios used for the autopilot because we didn't want to have dependencies between the two systems.
 
@@
 The PM5's NTSC video signal outputs a 480 line video of the current ``viewfinder'', the live video that is also displayed on the rear LCD screen. 
-Still pictures are taken based on targets spotted in this video feed. Those images are stored on the camera's memory card.
+Still pictures are taken based on targets spotted in this video feed and are stored on the camera's memory card.
+
+Each still picture is 5 megapixels, with an actual resolution of 2592 x 1944.
+To determine how the images will appear from different altitudes, we have a table that maps the size of features in the world to pixels in our images at different altitudes.
+We also see how this mapping will be affected when the camera is pointed straight down compared to when it is tilted at a 45 degree angle.
+We can see that at 100 ft.
+each pixel will represent a little over $1/3$ of an inch when looking straight down, which should be a sufficient level of detail to accurately identify any target.
+However, at 500 ft.
+each pixel is 2 inches wide when looking straight down, which will lead to blurry images.
+We are confident that this level of detail will be enough to identify the shapes and colors of all targets at any required altitude, but we may have some difficulty identifying alpha-numerics on very small targets at high altitudes.
+As a first year team, we feel that this is an acceptable level of detail.
+
+\begin{tabular}{ccc}
+{\bf altitude (ft)} & {\bf pixel width at $0,^{\circ}$ (in)} & {\bf pixel width at $45,^{\circ}$ (in)} \\
+       100 &       0.38 &       0.54 \\
+       200 &       0.76 &       1.08 \\
+       300 &       1.14 &       1.62 \\
+       400 &       1.52 &       2.16 \\
+       500 &       1.90 &       2.69 \\
+\end{tabular}
 
@@
 Please also give the resolving power of the video viewfinder
-
-\subsection{Camera-Computer Interface}
-
-We access the still pictures on the memory card via the PM5's USB connector.
-When attached to a computer, the PM5 powers the camera off and appears as a USB Mass Storage Device.
-In order to use the PM5 as both a camera and Mass Storage Device in flight, we have spliced a transistor into the power line on the USB cable.
-We control whether the camera detects it is connected via USB by blocking or allowing current in that line.
-
-% cogan, please finish up this section. Some of the information may be redundant.
-% we also need a figure of this system, labeling all the components and connections.
-
-% from michael varga:
-%Overview of Ground Station Commands
-%Some of the commands that the ground station can issue to the plane are capture image, zoom in, zoom out, crop image, download crop, generate crop, pan left, pan right, tilt up and tilt down.
-These commands are initiate by a simple button press on the ground station gui program.
-The ground station sends a signal to the plane computer through a radio.
-The plane software then handles the command directly or may forward it on to the microcontroller.
-The capture image, pan and tilt commands are forwarded onto the microcontroller as they involve the lanc interface with the camera and the servos.
-The download crop and generate crop commands are handled directly by the computer since the data necessary to carry out these functions exists on its hard drive..
-%Target Verification
-%A live video feed is forwarded from the camera through a ntsc video transmitter.
-The ground station laptop contains a ntsc video receiver so it may be displayed on the screen.
-When the operator sees a target in the view of the camera he issues a command to the plane to capture an image of the camera’s current view.
-These images will be stored on the plane and ready for download when instructed by the ground station operator.
-%The original full resolution images are too large to be downloaded directly.
-A lower resolution picture is generated for every high resolution image that is captured.
-The ground station operator must download and look at the thumbnail to select a rectangular region that encompasses the target.
-A command is issued to the plane that generates a crop from the original higher resolution image using the selected rectangular region.
-The ground station operator may then download this newly generated crop to verify the target with more precision.
 
@@
 When a picture is taken on the plane it is immediately saved on the camera's memory card.
-In order to view that picture on the ground it must first be transferred to the plane's main computer (beagle board), and then downloaded to the Image Station over the XBee connection.
+In order to view that picture on the ground it must first be transferred to the plane's main computer (beagle board).
+This is done via the PM5's USB connector.
+When attached to a computer, the PM5 powers the camera off and appears as a USB Mass Storage Device.
+In order to use the PM5 as both a camera and Mass Storage Device in flight, we have spliced a transistor into the power line on the USB cable.
+This controls whether the camera detects it is connected via USB by blocking or allowing current in that line.
 
@@
 Once images are on the plane's computer they are ready to be downloaded over a 2.4GHz XBee radio link.
-Each image is fairly large, and our bandwidth is only around 115200 baud % actually throughput, at worst-case, is half this.
-, so we select only the most important part of the images to download.
-The entire image is first downloaded as a low-res `thumbnail.'  Once that is available for viewing, the operator can select an area of interest and download a high-res crop of just that area.
-This allows us to quickly obtain high resolution images of the targets of interest.
+Each picture taken by the camera is about 500KB in size, while our theoretical maximum throughput is only 115200 baud (and only about half of that in practice).
+Each full image takes about a minute to download, so we created a method to speed up the process.
+The entire image is first downloaded as an 800x600 resolution `thumbnail.'  Once the thumbnail is available for viewing, the operator can select an area of interest and download a high-res crop of just that area.
+Downloading both the crop and the thumbnail generally takes under 30 seconds, less than half the time of downloading the full image, allowing us to quickly obtain high resolution images of the targets of interest.
 
-% Please state the sizes of the thumbnail and cropped images, compared to the total image size, and state the average download times of each, in a table. this will make our improvements dramatic.
+\begin{tabular}{rrr}
+ & {\bf Size (KB)} & {\bf Download Time (s)} \\
+ {\bf Original Image} &        500 &       69.4 \\
+ {\bf Thumbnail} &        125 &       17.4 \\
+ {\bf Crop} &         25 &        3.5 \\
+\end{tabular}
+
+Below a diagram is included that shows the entire flow of events from start to finish in our Imaging System for capturing an image and then downloading it.
+
+\begin{figure} [H]
+  \centering
+  	\includegraphics[width=0.9\textwidth]{../images/CaptureProcess.jpg}
+  	\caption[Image Capture Process]{Flow of events for capturing an image}
+  	\label{fig:imagecaptureprocess}
+\end{figure}
+
+\begin{enumerate}
+\item Command to take a picture is sent to Plane from Image Station
+\item Plane sends command to camera, picture is taken and acknowledgement is sent to Image Station
+\item Operator issues command to make images available for download
+\item Plane acknowledges request and begins transfer of images from internal camera memory to the BeagleBoard.  Sends notification when transfer is complete.
+\item Newly available image(s) are automatically put on a queue to be downloaded.
+\item A request to download a 250 byte chunk of the image at the top of the queue is sent to the plane.
+\item The plane sends down the requested chunk of image.
+\item Steps 6-7 are repeated throughout the duration of the flight.
+These image downloading commands are low priority and will only be sent if there are no other commands to be executed, This allows for high priority commands like taking a picture or moving the camera to be executed at any time.
+\end{enumerate}
+
+\subsection{Identifying target characteristics}
+
+When an image has been fully downloaded to the ground targets can be identified and tagged with metadata.
+Information about the plane's orientation, gps coordinates, altitude, yaw/pitch/roll, and pan/tilt of the camera are sent down with the image.
+Using this information the GPS location of the target can be determined using the intrinsic properties of the camera and some basic trigonometry.
+Target shape, color, alphanumeric, alphanumeric color, and orientation are all entered manually by the operator.
 
@@
 \end{figure}
-
-\subsection{Identifying target characteristics}
-
-When an image has been fully downloaded to the ground targets can be identified and tagged with metadata.
-Information about the plane's orientation, gps coordinates, altitude, yaw/pitch/roll, and pan/tilt of the camera are sent down with the image.
-Using this information the GPS location of the target can be determined using the intrinsic properties of the camera and some basic trigonometry.
-Target shape, color, alphanumeric, alphanumeric color, and orientation are all entered manually by the operator.
 

### 2010-05-25 paper/paper.tex
reorganized electronics system engineering a little bit, and better described single board linux computer

@@
 The biggest system design choice we made was to use a single board Linux-based flight computer as the platform for both systems.
-We felt that the best way to develop software which interfaces with a wide variety of analog and digital electrical systems, has complicated information flow requirements,
+We felt that the best way to develop software which interfaces with a wide variety of analog and digital electrical systems, and has complicated information flow requirements.
 
-We felt the Paparazzi system was full featured and proven, but we weren't satisified with the capabilities of the existing Paparazzi hardware. The ``Tiny'' v2.1\cite{paparazzi_tinyv21}, the latest
+We had sketched plans for an electronics system built entirely on a Linux-based flight computer as early as July 2009, but planned on using a more standard microcontroller-based autopilot system with an auxillary Linux-based computer for the imaging system until December 2009.
+After burning out our microcontroller based autopilot for the second time, we were faced with a long lead time for an expensive replacement part.
+On that impetus, we redesigned our electronics systems to eliminate the microcontroller autopilot and ensure all of our replacment parts would be inexpensive and commercially available from many vendors.
+The move to a single Linux-based flight computer, while not without obstacles, was a successful design decision, as we'll describe in this paper.
+
+\subsubsection{Autopilot System Engineering}
+
+We felt the Paparazzi system was full featured and proven, but we weren't satisified with the capabilities of the existing Paparazzi hardware.
+The ``Tiny'' v2.1\cite{paparazzi_tinyv21}, the latest autopilot hardware released by the Paparazzi project, is based on a 64 pin ARM microcontroller.
+Almost all of its input-output capabilities are dedicated to autopilot functions, making it difficult to expand functionality
+
+ The first task was to select hardware which would allow us to port the airborne code from the Paparazzi's standard microcontroller-based autopilot hardware to our flight computer.
 
@@
 Along with careful preflight testing, careful observation, and definitive analysis of any errors encountered during flight, these measures meet the spirit and letter of competition rules.
-
-\subsubsection{Autopilot System Engineering}
-
-As stated previously, the Paparazzi system is the most capable and proven open-source autopilot system available.
-The first task was to select hardware which would allow us to port the airborne code from the Paparazzi's standard microcontroller-based autopilot hardware to our flight computer.
 

### 2010-05-25 paper/paper.tex
finished up text of electronics, autopilot system engineering

@@
 The ``Tiny'' v2.1\cite{paparazzi_tinyv21}, the latest autopilot hardware released by the Paparazzi project, is based on a 64 pin ARM microcontroller.
-Almost all of its input-output capabilities are dedicated to autopilot functions, making it difficult to expand functionality
+Almost all of its input-output capabilities are dedicated to autopilot functions, making it difficult to expand functionality and share information (for instance, position and orientation estimates) with payload components.
+As previously stated, the Tiny is expensive, fragile\footnote{In our own admittedly clumsy experience} and often subject to long lead times.
 
- The first task was to select hardware which would allow us to port the airborne code from the Paparazzi's standard microcontroller-based autopilot hardware to our flight computer.
+In order to replace the functionality of the Tiny with a flight computer, we needed to select hardware which could provide the flight control software with information from the attitude sensor, GPS module, and long distance telemetry radio link.
+We also needed a servo output interface.
+The most important feature provided by the Tiny was the real-time processing and multiplexing of the safety pilot's radio control signals with the autopilot control signals.
+The flight computer's servo output hardware must manage this multiplexing functionality in a way that is not tied to the function of the flight computer itself.
 
@@
 Because our entry is essentially an experimental system, and will always be operated with a safety pilot present, we engineered the safety pilot override to operate independently from the flight computer.
-We have ensured that, short of losing power to the radio control receiver and servos, the safety pilot will always be able to take over at a moment's notice.
+We have ensured that, short of losing power to the radio control receiver and servos, the safety pilot will always be able to take over with the flick of a single switch on his radio control transmitter.
 We have also ensured that the flight computer's servo controller will be able to detect a loss of computer control and servos to a recoverable default.

### 2010-05-25 paper/paper.tex
changed pantilt dimensions in text based on new measurments, new figure 6b
cogans changes

--- 2010-05-25 paper/paper.tex
@@
-\documentclass[10pt,twocolumns]{report}
+\documentclass[10pt]{report}
 \usepackage[pdftex]{graphicx}
-\usepackage[left=1.5in, right=1.5in, top=1.25in, bottom=1.25in]{geometry}
 \usepackage{hyperref}
@@
 \usepackage{subfig}
-\usepackage{wrapfig}
 \usepackage{listings}
@@
  Bill Perkowski,
-and  Mike Korb, 
-%and Chris (test pilot) 
-amongst many who spent time mentoring us.
+ Mike Korb,
+ (the man who donated daedalus), and
+ Chris (test pilot) amongst many who spent time mentoring us.
 
@@
 We would like to acknowledge
-Louis Stumpf, for donating the aircraft which became the \emph{Daedalus};
 Jeff Perkins, for his prolific networking and promotion of the team;
-Gregg Peters of BP Hobbies for getting us out of numerous binds;
+Gregg (last name) of BP Hobbies for getting us out of numerous binds;
 Dave Devries and Michael Maia, of Invensense, for their technical guidance;
@@
 
-\begin{figure}[b!]
-	\centering
-	\includegraphics[width=0.6\textwidth]{../images/wreckage.jpg}
-	\caption{The recovery of our 2009 entry, the \emph{Icarus}}
-	\label{fig:wreckage}
-\end{figure}
+\section{Systems Engineering}
 
-\newpage
-\section{Systems Engineering}
 \subsection{Goals}
@@
 We also wanted to balance building with buying, keeping in mind our slim finances.
-While we couldn't build every part of our entry from scratch, we tried to build whatever we could given the 16th June competition deadline.
+While its impossible to truly start from scratch\footnote{``If you want to make an apple pie from scratch, you must first create the universe'' - Dr.
+Carl Sagan \cite{saganquote}}, we tried to build whatever we could given the 16th June competition deadline.
 
-%\footnote{``If you want to make an apple pie from scratch, you must first create the universe'' - Dr. Carl Sagan \cite{saganquote}},
+% footnote is broken (breaks from bottom of page one onto bottom of page two), something about the documentclass needs to be changed so it doesn't assume pages are next to each other (book-like)?
 
@@
 
-Our team was divided by area of expertise into a mechanical team and an electrical team.The mechanical team began the year with the goal of producing two competition worthy airframes.
-These were to become the \emph{Daedalus} and \emph{Knight Two}.
-
-The teams made design decisions based on the aforementioned competition goals.
-
-One of the most difficult design decisions was to quantify ``ample payload.'' Payload size and weight is the major determining factor in aircraft size.
-We originally assumed a 2.5lb electronics (computer + batteries) payload and a 2.5lb pan-tilt camera 8'' in diameter, requiring 6'' of vertical space extruding from the bottom of the aircraft.
-These considerations introduced two major design parameters: the aircraft must be capable of carrying a minimum 5lb payload, and must be situated to provide clearance for the pan tilt unit.
-
-Additionally, we sought an easily controllable, stable platform to accommodate the autopilot, and an ability to fly at low speeds, in order to more easily capture clear images.
+Our team was divided by area of expertise into a mechanical team and an electrical team.
+The mechanical team began the year with the goal of producing two competition worthy airframes. These were to become the \emph{Daedalus} and \emph{Knight Two}.
 
 We determined, after consulting with R/C experts, that a high wing trainer is the most stable and proven fixed-wing design.
-Trainers traditionally have light wing loading for slow speed flight, and are able to carry a comparatively large payload.
-Additionally, a large plane with lower wing loading is easier to control, and easier to recover if power is lost.
-Consciously agreeing to err on the side of caution, we sought a large plane capable of accommodating a worst-case scenario payload, and settled on a wingspan on the order of 10'-12'.
-Though lighter and smaller payload alternatives were found as the year progressed, these merely leave us with more freedom for future modifications.
-%That last sentence sounds kind of eh, I'm stuck on a decent wording, so I'm leaving it for now.
+Trainers traditionally have light wing loading for slow speed flight, and will be able to carry payload
 
-The decision to employ a large, high wing trainer, however, forced us to rethink our original pan-tilt design.
-Last year, in order to accommodate the pan-tilt unit, we designed a pusher plane, enabling us to mount the pan-tilt under the plane's nose, which had sufficient clearance.
-Since trainers are generally designed with front-mounted engines (I presume?), we were forced to engineer a new solution requiring significantly less clearance.
-In the long run, we modified the pan-tilt mount to allow us to recess more of the pan-tilt unit into the body of the plane---this had the added benefit of providing more protection to an expensive piece of equipment on takeoff and landing.
+One of the most difficult design decisions was to quantify ``ample payload.'' The payload size and weight is the major determining factor in the aircraft size.
+% -- amanda will finish this by mon noon: 
+We started off assuming an electronics (computer + batteries) payload of 2.5lbs and a pan-tilt camera of 2.5lbs and 8 inch diameter.
 
@@
 
-We felt the Paparazzi system was full featured and proven, but we weren't satisified with the capabilities of the existing Paparazzi hardware. The ``Tiny'' v2.1\cite{paparazzi_tinyv21}, the latest
+We felt the Paparazzi system was full featured and proven, and we wanted to adapt it to our needs.
 
@@
 \subsubsection{Daedalus}
-\begin{wrapfigure}{l}{0.4\textwidth}
-	\includegraphics[width=0.37\textwidth]{../images/daedalus_isometric.jpg}
-	\caption{The \emph{Daedalus}}
-	\label{fig:daedalus}
-\end{wrapfigure}
+
 A member of our local R/C club, Tri-County RC \cite{tricountyRC}, donated a ten-foot wingspan high wing trainer to our club.
@@
 
- \begin{wrapfigure}{r}{0.4\textwidth}
-	\includegraphics[width=0.37\textwidth]{../images/daedalus_sandbagtest.jpg}
-	\caption{Load testing the underside of the \emph{Daedalus}'s wings. Shown with 16lbs of load.}
-	\label{fig:sandbagtest}
-\end{wrapfigure}
 The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symmetrically and simultaneously on the underside of each wing, starting from the innermost rib.
@@
 
-\begin{wrapfigure}{l}{0.4\textwidth}
-	\centering
-	\subfloat[The unfinished \emph{Knight Two}]{\label{fig:knighttwo_construction} 
-			\includegraphics[width=0.37\textwidth]{../images/knighttwo_uncovered_outdoors.jpg}}
-\\
-	\subfloat[A completed 12 Foot Telemaster. Photo Credit: Hobby Lobby] {\label{fig:knighttwo_12ft_telemaster} 
-			\includegraphics[width=0.37\textwidth]{../images/telemaster_12ft_hobbylobby.jpg}}
-  	\label{fig:knighttwo}
-  	\caption[Knight Two]{Knight Two}
+Figure goes here: Photos of Daedalus in construction, sandbag testing
 
-\end{wrapfigure}
 Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations
+
+Figure may go here: Line drawing of Daedalus. JOHN
 
@@
 
+\begin{figure}
+	\centering
+	\subfloat[The unfinished \emph{Knight Two}]{\label{fig:knighttwo_construction} 
+			\includegraphics[width=0.4\textwidth]{../images/knighttwo_underconstruction.jpg}}
+	\subfloat[A completed 12 Foot Telemaster. Photo Credit: Hobby Lobby] {\label{fig:knighttwo_12ft_telemaster} 
+			\includegraphics[width=0.4\textwidth]{../images/telemaster_12ft_hobbylobby.jpg}}
+  	\label{fig:knighttwo}
+  	\caption[Knight Two]{Knight Two}
+\end{figure}
 Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations
@@
 The pan servo is a Hitech HS 785HB \ref{hitechhs785}, and is capable of \degrees{1260} rotation (3.5 revolutions).
-Both servos are powered and controlled by the Pan-tilt control board.
+Both servos are powered and controlled by the Camera Control board.
 
@@
 
-\subsection{Safety Observers}
-A safety observer is defined as a person, devoid of any additional specific responsibilities, who remains in the starting area during all pre-flight procedures.
-The safety observer is expected to watch for unsafe conditions, included but not limited to non-essential personnel in the starting, takeoff, and/or flight areas; loose mechanical connections; failed surface and/or range checks; and other forms of mechanical failure.
-The safety observer has the absolute authority to give the order to kill the ignition and abort startup immediately upon observing any unsafe condition.
-The safety observer is responsible for communicating possible conditions to the startup team as they arise.
-At minimum, one safety observer is required to start the engine, though we typically have more.
-
 \newpage
@@
 		\end{itemize}
-	\item Position wingmen\footnote{Two men, one positioned at each wing tip, preventing the plane from moving during engine start and subsequent surface checks.}, centerman\footnote{Man straddling the fuse, preventing the plane from moving during engine start and subsequent surface checks}.
+	\item Position wingmen\footnotemark, centerman\footnotemark.
 	\item Pilot turns on transmitter, announces ``transmitter on.''

### 2010-05-25 paper/paper.tex
added two software diagram figures, and the photo of the pantilt

--- 2010-05-25 paper/paper.tex
@@
-\documentclass[10pt,twocolumns]{report}
+\documentclass[10pt]{report}
 \usepackage[pdftex]{graphicx}
@@
 \subsection{Hardware}
-Single board computer: beagleboard.
+Description of components selected.
+Description of each input, output interface.
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\textwidth]{../images/hardware_conn_diagram.pdf}
+	\caption{Hardware connection diagram for the autopilot and imaging systems. Only one of the 9 axis AHRS and Thermopile Attitude Sensor are used.}
+	\label{fig:hardware_conn_diagram}
+\end{figure}
 
-Description of each input, output interface.
-
-Figure goes here: Hardware connection diagram. PAT
 \subsection{Software}
 \label{sec:autopilot_software}
-PAT
 
-Figure goes here: interprocess communication. PAT
+\begin{figure}
+	\centering
+	\includegraphics[width=0.9\textwidth]{../images/software_comm_diagram.pdf}
+	\caption{Communications diagram}	
+	\label{fig:software_comm_diagram}
+\end{figure}
+
 \subsection{Ground Station}
@@
 We are confident that this level of detail will be enough to identify the shapes and colors of all targets at any required altitude, but we may have some difficulty identifying alpha-numerics on very small targets at high altitudes.
-As a first year team, we feel that this is an acceptable level of detail.
+As a first year team, we feel that meeting the ``threshold'' rather than the ``objective'' goals, at the edge of the altitude envelope, is an acceptable level of performance.
 
 \begin{tabular}{ccc}
-{\bf altitude (ft)} & {\bf pixel width at $0,^{\circ}$ (in)} & {\bf pixel width at $45,^{\circ}$ (in)} \\
+\\{\bf altitude (ft)} & {\bf pixel width at $0,^{\circ}$ (in)} & {\bf pixel width at $45,^{\circ}$ (in)} \\
        100 &       0.38 &       0.54 \\
@@
 	\subfloat[Isometric View]{\label{fig:pantilt_isometric} 
-			\includegraphics[width=0.75\textwidth]{../images/pantilt_isometric.png}}
+			\includegraphics[width=0.45\textwidth]{../images/pantilt_isometric.png}}
+	\subfloat[Side View.]{\label{fig:pantilt_side} 
+			\includegraphics[width=0.45\textwidth]{../images/pantilt_side.png}}
 \\
-	\subfloat[Side View.]{\label{fig:pantilt_side} 
-			\includegraphics[width=0.75\textwidth]{../images/pantilt_side.png}}
+	\subfloat[Photograph of pan tilt unit installed in \emph{Daedalus}]{\label{fig:pantilt_photo}
+			\includegraphics[width=0.75\textwidth]{../images/pantilt_photo1.jpg}}
 	\caption{Pan Tilt Unit}
@@
 
-The total height of the imaging system is 8.75", only 4'' of which is exposed below the fuselage of the aircraft.
+The total height of the imaging system is 6.75", a maximum 2'' of which is exposed below the fuselage of the aircraft.
 The rotating assembly is, on its widest dimension 2.13", and would fit inside a 2.5" cylinder.

### 2010-05-25 paper/paper.tex
many many edits to the introduction through the end of airframe system engineering

@@
 
-The Rutgers Autonomous Aircraft Team was formed in September 2008 with the goal of entering the 2009 AUVSI UAS competition.
-We didn't make it to the 2009 competition after loosing our aircraft only four weeks before.
+The Rutgers Autonomous Aircraft Team was formed in September 2008 with the goal of entering the 2009 Association for Unmanned Vehicle Systems International (AUVSI)\cite{auvsiweb} Unmanned Air Systems (UAS) competition\cite{uasweb}.
+We didn't make it to the 2009 competition after losing our aircraft only four weeks before.
 Our imaging system and autopilot system were also far from ready.
-
-In terms of NCAA sport eligibility, one could consider 2009 our `red shirt' year. In 2010 we are still rookies.
+In terms of NCAA sport eligibility, one could consider 2009 our `red shirt' year.
+In 2010 we are still rookies.
 
@@
 Rutgers Engineering Governance Council,
-Rutgers Alumni Association, and
-WINLAB, a laboratory of the Rutgers University Department of Electrical Engineering.
-% What was Chris Brocco's Uncle's Company? Something Automation?
-BP Hobbies,
+Rutgers Alumni Association,
+WINLAB, a laboratory of the Rutgers University Department of Electrical Engineering, 
+and BP Hobbies.
 
 We would like to acknowledge
-Louis Stumpf, for donating the aircraft which became the \emph{Daedalus};
+Louis Stumpf, for donating the aircraft that became the \emph{Daedalus};
 Jeff Perkins, for his prolific networking and promotion of the team;
@@
 Professor Chung-chieh Shan, for overseeing Pat's independent study on the IMU design;
-Ivan Seskar, for keeping us (and many other undergrads) in the black;
+Ivan Seskar, for keeping us (and many other undergrads) in the black; and
 Roger Kondos and Brij Pathak, for contributing to the thermopile attitude sensor.
@@
 \subsection{Goals}
-For this entry, we took a close look at our goals and requirements, and dialed back our approach to ensure we would meet the most basic competition requirements successfully.
+For this entry we took a close look at our goals and requirements, and attempted to meet the ``threshold'' competition requirements successfully.
 We started the design by forming, as a team, some broad goals based on our experience in the 2009 competition.
-These formed the basis on which sub-teams and individuals tackled the smaller problems while still keeping an eye on the big picture.
+These goals allowed sub-teams and individuals to tackle the smaller problems while still keeping an eye on the big picture.
 
-We also wanted to balance building with buying, keeping in mind our slim finances.
-While we couldn't build every part of our entry from scratch, we tried to build whatever we could given the 16th June competition deadline.
+We also wanted to balance building components with buying components, keeping in mind our slim finances.
+While we couldn't build every part of our entry from scratch, we were motivated to build as much of our systems as possible for the learning experience.
+We tried to buy only components which we didn't have the time or resources to build ourselves.
 
@@
 	\item The imaging system must be simple and capable.
-We assume the operator will be burdened with guiding the camera and identifying targets until time and resources permit us to add sensor information and control loops for guidance, and computer vision for identification.
-It will be custom built, because we could not find any affordable systems available commercially.
+We assume the operator will be burdened with guiding the camera and identifying targets manually.
+We couldn't guarantee time and resources would permit us to add sensor information and control loops for camera guidance and computer vision for target identification.
 \end{itemize}
 
-We made design decisions of which competition goals we will attempt. 
-We will be performing a manual take off, and giving the autopilot system control of the aircraft once the plane is in the air.
+We will be performing a manual take off and giving the autopilot system control of the aircraft once the plane is in the air.
 We will also perform a manual landing.
-We decided not to pursue the reach goals of autonomous takeoff and landing based on our inexperience and the risk, during development and competition, to our aircraft.
+We decided not to pursue the reach goals of autonomous takeoff and landing based on our inexperience and the risk of a crash during development.
 \subsection{System Design Decisions}
@@
 
-The teams made design decisions based on the aforementioned competition goals.
-
 One of the most difficult design decisions was to quantify ``ample payload.'' Payload size and weight is the major determining factor in aircraft size.
-We originally assumed a 2.5lb electronics (computer + batteries) payload and a 2.5lb pan-tilt camera 8'' in diameter, requiring 6'' of vertical space extruding from the bottom of the aircraft.
+We originally assumed a 2.5lb electronics (flight computer, sensors, and batteries) payload and a 2.5lb pan-tilt camera 8'' in diameter, requiring 6'' of vertical space below the bottom of the aircraft.
 These considerations introduced two major design parameters: the aircraft must be capable of carrying a minimum 5lb payload, and must be situated to provide clearance for the pan tilt unit.
 
-Additionally, we sought an easily controllable, stable platform to accommodate the autopilot, and an ability to fly at low speeds, in order to more easily capture clear images.
+We also needed an easily controllable, stable platform to accommodate the autopilot, and we needed to keep flight speed low to make image capture as easy as possible.
+After consulting with R/C experts, a high wing trainer is looked like the most stable and proven fixed-wing design.
+Trainers traditionally have light wing loading for slow speed flight, and are able to carry a comparatively large payload.Their lower wing loading is easier to control, and easier to recover if power is lost.
 
-We determined, after consulting with R/C experts, that a high wing trainer is the most stable and proven fixed-wing design.
-Trainers traditionally have light wing loading for slow speed flight, and are able to carry a comparatively large payload.
-Additionally, a large plane with lower wing loading is easier to control, and easier to recover if power is lost.
-Consciously agreeing to err on the side of caution, we sought a large plane capable of accommodating a worst-case scenario payload, and settled on a wingspan on the order of 10'-12'.
-Though lighter and smaller payload alternatives were found as the year progressed, these merely leave us with more freedom for future modifications.
-%That last sentence sounds kind of eh, I'm stuck on a decent wording, so I'm leaving it for now.
+Erring on the side of caution, we decided our aircraft should be capable of accommodating a worst-case scenario payload. 
+We were, in part, cautious due to a previous error where we rebuilt an entire aircraft to accommodate a larger payload.
+We decided on a high wing trainer with a 10' to 12' wingspan, so that we could accommodate up to 5lbs of payload while maintaining suitable wing loading.
+Though we managed to keep our payload weight well under 5lbs, the only major drawback to having an oversize aircraft is the difficulty of transport.
 
-The decision to employ a large, high wing trainer, however, forced us to rethink our original pan-tilt design.
-Last year, in order to accommodate the pan-tilt unit, we designed a pusher plane, enabling us to mount the pan-tilt under the plane's nose, which had sufficient clearance.
-Since trainers are generally designed with front-mounted engines (I presume?), we were forced to engineer a new solution requiring significantly less clearance.
-In the long run, we modified the pan-tilt mount to allow us to recess more of the pan-tilt unit into the body of the plane---this had the added benefit of providing more protection to an expensive piece of equipment on takeoff and landing.
+Our selection of a trainer type aircraft encouraged us to rethink our pan tilt design.
+We based our initial assumptions of an 8'' diameter pan and tilt unit, mounted underneath the aircraft, based on a design that was to be mounted on the nose of our 2009 entry, an aircraft with a pusher motor configuration.
+When we considered the possibilities of either dramatically increasing the ride height of our aircraft, or making significant modifications to the width of the fuselage, we decided to redesign our pan tilt system to fit in a more suitable envelope.
 

### 2010-05-25 paper/paper.tex
redid imaging system engineering a bit

@@
 
-We had sketched plans for an electronics system built entirely on a Linux-based flight computer as early as July 2009, but planned on using a more standard microcontroller-based autopilot system with an auxillary Linux-based computer for the imaging system until December 2009.
+We had sketched plans for an electronics system built entirely on a Linux-based flight computer as early as July 2009, but planned on using a more standard microcontroller-based autopilot system with an auxiliary Linux-based computer for the imaging system until December 2009.
 After burning out our microcontroller based autopilot for the second time, we were faced with a long lead time for an expensive replacement part.
-On that impetus, we redesigned our electronics systems to eliminate the microcontroller autopilot and ensure all of our replacment parts would be inexpensive and commercially available from many vendors.
+On that impetus, we redesigned our electronics systems to eliminate the microcontroller autopilot and ensure all of our replacement parts would be inexpensive and commercially available from many vendors.
 The move to a single Linux-based flight computer, while not without obstacles, was a successful design decision, as we'll describe in this paper.
@@
 
-We felt the Paparazzi system was full featured and proven, but we weren't satisified with the capabilities of the existing Paparazzi hardware.
+We felt the Paparazzi system was full featured and proven, but we weren't satisfied with the capabilities of the existing Paparazzi hardware.
 The ``Tiny'' v2.1\cite{paparazzi_tinyv21}, the latest autopilot hardware released by the Paparazzi project, is based on a 64 pin ARM microcontroller.
@@
 Both Pat Hickey and Bradley Lord became intimately familiar with the source code and control flow inside the autopilot software, and all patches were reviewed and tested by both engineers.
-Numerous other safety measures will be described in Section \ref{sec:autopilot_software}.
+Numerous other safety measures are described in Section \ref{sec:autopilot_software}.
 
-Because our entry is essentially an experimental system, and will always be operated with a safety pilot present, we engineered the safety pilot override to operate independently from the flight computer.
+Because our entry is essentially an experimental system and will always be operated with a safety pilot present, we engineered the safety pilot override to operate independently from the flight computer.
 We have ensured that, short of losing power to the radio control receiver and servos, the safety pilot will always be able to take over with the flick of a single switch on his radio control transmitter.
@@
 
-When designing the imaging system our major goals were to have 1) a camera with XXX range of motion.
-2) real-time video feed from the plane.
-3) ability to control the camera in real-time and take pictures.
-4) ability to download pictures while in flight.
-For the camera we chose the Sony Bloggie \cite{bloggie} because it was small, light, cheap, and took high resolution images and video.
-In addition it came with a built in rotating lens, which simplified the construcion of the pan-tilt unit, and it can also be controlled remotely over usb, so it interfaced with our flight linux computer.
-To communicate with the plane we developed custom ground station software to meet our needs and installed two sets of radios, one for the video feed, and another for communicating with the plane and downloading images.
-These radios are independent from the set of radios used for the autopilot because we didn't want to have dependencies between the two systems.
+When designing the imaging system our major goals were to have 
+1) a camera able to capture imagery ``up to \degrees{60} in all directions from vertically below the aircraft''\cite{competition_rules}, as per the rules; 
+2) a real-time video feed;
+4) the ability to download pictures while in flight; and
+3) the ability to remotely control the camera, particularly to trigger pictures.
+
+It was difficult to find a camera to base our imaging system on which met all of these requirements.
+Many still cameras can be fitted on a pan tilt system to requirement 1, and provide a real-time ``viewfinder'' video feed to meet requirement 2.
+Requirement 3 could typically be managed by attaching the camera to a single board computer by USB cable, though usually the camera powers down when attached to the computer, and some sort of switch is needed to deactivate the USB cable once the data transfer is complete.
+
+Requirement 4 had the greatest effect on our design decision.
+Typically, only high-end SLR-type still cameras provide a socket for remote control.
+These are beyond our price range, and are much larger and heavier than we would like.
+We built a prototype using a Canon Powershot S5 IS \cite{powershotS5}, a camera which has no electrical remote control input, by soldering wires to the push-button pads for the shutter and zoom buttons.
+Those wires were attached to isolating transistors, and a microcontroller was used to control the buttons remotely But after some time, this clever circumvention burned out the whole camera, and we abandoned the scheme.
 

### 2010-05-25 paper/paper.tex
missing word

@@
 We built a prototype using a Canon Powershot S5 IS \cite{powershotS5}, a camera which has no electrical remote control input, by soldering wires to the push-button pads for the shutter and zoom buttons.
-Those wires were attached to isolating transistors, and a microcontroller was used to control the buttons remotely But after some time, this clever circumvention burned out the whole camera, and we abandoned the scheme.
+Those wires were attached to isolating transistors, and a microcontroller was used to control the buttons remotely But after some time, this clever circumvention accidentally burned out the whole camera, and we abandoned the scheme.
 

### 2010-05-25 paper/paper.tex
finished edits up to the end of Knight Two subsubsection

@@
 At the time of this writing, we have completed and flown a single flight vehicle, which we have named the \emph{Daedalus}.
-We have also nearly completed a very similar backup airframe which serve as a backup in case the \emph{Daedalus} has a disaster.
+We have also nearly completed a very similar backup airframe which will serve as a backup in case the \emph{Daedalus} has a disaster.
 The \emph{Knight Two} will be fitted with nearly identical R/C flight electronics, and accept the same autopilot and imaging system payload as the \emph{Daedalus}.
@@
 \subsection{Airframe}
-
-We elected to pursue a kit-type airplane rather than our own design. 
-Our own custom design, the \emph{Icarus}, suffered a catastrophic structural failure last year.
-From this, we learned that custom structures require much more testing and revision than a proven kit.
-We also reconsidered our payload requirements and decided we could easily modify a kit plane to accept our autopilot and imaging system.
 
@@
 \end{wrapfigure}
-A member of our local R/C club, Tri-County RC \cite{tricountyRC}, donated a ten-foot wingspan high wing trainer to our club.
+Louis Stumpf, a member of our local R/C club, donated a 10ft 1in (3.07m) wingspan, 7ft 8in (2.33m) long, high wing trainer to our club. It met all of the requirements outlined previously.
+
 The plane, first built out of foam and balsa from long-lost plans in 1986, was accepted graciously, but required quite a bit of work before it was once again flight-worthy.
@@
 We stripped and rebuilt nearly the entire airframe, adding carbon-fiber reinforcements to the wing in anticipation of increased wing loading due to our payload.
-Once rebuilt and covered, we tested the \emph{Daedalus} weighed 19.5 lbs with engine.
+Once rebuilt and covered,  the \emph{Daedalus} weighed 19.5 lbs with engine.
 A full 24oz fuel tank, plus 3.5 lbs electronics payload, would bring the maximum takeoff weight to 25.5lbs.
@@
 We wanted to certify the wings and wing root would withstand large dynamic forces in flight. 
-We designed a test similar to mimic the 
+We designed a test to mimic the 
 Federal Airworthiness Standards' structural requirements\footnote{Special Federal Aviation Regulation No.
@@
 
-Satisfied, made engine and surface tests (see \ref{sec:appendix_preflight}) before taxiing the plane around a parking lot.
+Satisfied, we made engine and surface tests (see \ref{sec:appendix_preflight}) before taxiing the plane around a parking lot.
 We brought the plane back inside, examined it for signs of stress and, finding everything in good condition, deemed the \emph{Daedalus} flight-worthy.
@@
 We selected an Aero-Design 12 Foot Telemaster \cite{aerodesign} because it was available as a balsa kit and similar in size to the \emph{Daedalus}.
+The 12 Foot Telemaster has a 12ft 1in (3.68m) wingspan, and is 7ft 10in (2.38m) in length.We named it the \emph{Knight Two} after the Rutgers University mascot, the Scarlet Knight.
 
-We named it the \emph{Knight Two} after the Rutgers University mascot, the Scarlet Knight.
-We may rename it if someone can think of a better name.
-Pat needed to think up a decent name real fast in order to make this rough draft.
-
-Since our testing plan for the \emph{Daedalus} was successful, we will follow the same plan when, in the coming weeks, we finish building \emph{Knight Two}.
-
-Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations
-
-Figure may go here: Line drawing of Knight Two. JOHN
+At the time of this writing, the \emph{Knight Two} is 85\% complete. We plan to complete the aircraft and bring it to the competition as a spare.  
+Since our testing plan for the \emph{Daedalus} was successful, we will follow the same plan when the \emph{Knight Two} is completed.
 

### 2010-05-25 paper/paper.tex
another little rework of the canon s5 para

@@
 These are beyond our price range, and are much larger and heavier than we would like.
-We built a prototype using a Canon Powershot S5 IS \cite{powershotS5}, a camera which has no electrical remote control input, by soldering wires to the push-button pads for the shutter and zoom buttons.
-Those wires were attached to isolating transistors, and a microcontroller was used to control the buttons remotely But after some time, this clever circumvention accidentally burned out the whole camera, and we abandoned the scheme.
+
+We built a prototype using a Canon Powershot S5 IS \cite{powershotS5}, a camera which has no electrical remote control input.
+We soldered wires to the push-button pads for the shutter and zoom buttons so we could override them remotely.
+Those wires were attached to isolating transistors, and a microcontroller was used for control.
+After some time, this clever circumvention of the factory anti-static shielding resulted in a burnt out camera.
+We abandoned this scheme.
 

### 2010-05-25 paper/paper.tex

--- 2010-05-25 paper/paper.tex
@@
 
-\section{Introduction}
+\section{Introduction TWO}
 
@@
 We use a low voltage monitor in-line with the battery, which beeps when cell voltage falls below 7.2v.
-The autopilot electronics are powered off a completely separate battery for safety.
+The autopilot electronics are powered from a completely separate battery for safety reasons.
 
 \section{Autopilot}
-Our entry uses an autopilot system based off the open source 
-Paparazzi project\cite{paparazziweb}. 
+Our entry uses an autopilot system based off the open source Paparazzi project \cite{paparazziweb}.
 We ported the airborne code to Linux in order to use a single computer for all of our flight hardware and software.
@@
 \subsection{Camera}
-\begin{figure}
+
+\begin{figure} [h!]
 	\centering
@@
 We selected the Sony MHS-PM5 camera as the basis of our imaging system.
-Marketed as a pocket 1080i video camera capable of 5 megapixel (MP) still images, we primarily selected the camera because it features a swiveling lens at one end of the body and it features Sony's LANC (Local Application Control Bus System, ref{lanc}).
-Because the swiveling lens is at one end of the body, we can mount the camera primarily inside the fuselage of the aircraft, which will protect it from damage on takeoff and landing.
-Internal mounting also helps reduce drag.
-The swiveling lens is an important feature because it allows us to achieve two rotational degrees of freedom while only rotating the body about one axis.
-This simplifies the pan tilt unit design.
+Marketed as a pocket 1080i video camera capable of 5 megapixel still images, we primarily selected the camera because it features a swiveling lens at one end of the body and it features Sony's LANC (Local Application Control Bus System) \cite{lanc}.
+Because of the swiveling lens we were able to achieve two rotational degrees of freedom while only rotating the body of the camera about one axis.
+This greatly simiplified the pan-tilt design, and also allowed us to easily mount the camera primarily inside the fuselage of the aircraft, which protects it from damage during takeoff and landing, and reduces drag on the airframe.
 
 The LANC protocol is based on a 9600 baud UART serial connection, and is available on the PM5's 10-pin ``Multi-AV'' connector.
-Camera control commands, which include commands to zoom in, zoom out, and take a still picture, is publicly available \ref{boehmenl_lanc}.
+Camera control commands, which include commands to zoom in, zoom out, and take a still picture, are publicly documented.
 Also on the ``Multi-AV'' connector are the signals for NTSC video and audio output.
-We use both the NTSC output and LANC signals on this connector using a commercially available adapter cable \ref{lanc_multiav_cable}.
-
-% first citation should go to http://www.boehmel.de/lanc.htm
+We use both the NTSC output and LANC signals on this connector using a LANC Adapter cable \cite{lanc_adapter}.
 
@@
 Each still picture is 5 megapixels, with an actual resolution of 2592 x 1944.
-To determine how the images will appear from different altitudes, we have a table that maps the size of features in the world to pixels in our images at different altitudes.
-We also see how this mapping will be affected when the camera is pointed straight down compared to when it is tilted at a 45 degree angle.
-We can see that at 100 ft.
-each pixel will represent a little over $1/3$ of an inch when looking straight down, which should be a sufficient level of detail to accurately identify any target.
-However, at 500 ft.
-each pixel is 2 inches wide when looking straight down, which will lead to blurry images.
+To determine how the images will appear from different altitudes, we have created a table that maps the size of features in the world to pixels in our images at different altitudes.
+We also take into consideration the changes in distance to the ground when the camera is tilted at a \degrees(45) angle as opposed to vertically aligned with the plane.
+We can see that at an altitude of 100 ft.
+each pixel will represent an area a little over $1/3$ of an inch wide when looking straight down, which should be a sufficient level of detail to accurately identify any target.
+However, at an altitude of 500 ft.
+each pixel will represent an area almost 2 inches wide when looking straight down.
 We are confident that this level of detail will be enough to identify the shapes and colors of all targets at any required altitude, but we may have some difficulty identifying alpha-numerics on very small targets at high altitudes.
@@
 
+\begin{figure} [H]
+	\centering
 		\begin{tabular}{ccc}
-\\{\bf altitude (ft)} & {\bf pixel width at $0,^{\circ}$ (in)} & {\bf pixel width at $45,^{\circ}$ (in)} \\
+		\\{\bf altitude (ft)} & {\bf pixel width at 0$^{\circ}$ (in)} & {\bf pixel width at 45$^{\circ}$ (in)} \\
 	       100 &       0.38 &       0.54 \\
@@
 		\end{tabular}
-
-% please give details of still pictures, including the field of view per pixel.
-include a quick table of resolving power (projection of pixel field of view, center of frame, onto ground) at altitudes [100,200,300,400 agl] for straight down and 45 degree angle.
-Please also give the resolving power of the video viewfinder
+		\caption[Pixel Width Mapping]{mapping of altitude to the area each pixel will represent in an image}
+		\label{fig:pixelmapping}
+\end{figure}
 
@@
 
-In order to meet the target imaging requirments, we needed to add both pan and tilt actuators to the camera.
-The XX by YY degree lens field of view is insufficient to see off flight path targets, and would limit our capability in the target search area phase.
-We designed a pan tilt unit which make it possible to image the entire hemisphere below the fuselage of the aircraft.
+In order to meet the target imaging requirments we needed to add both pan and tilt actuators to the camera.
+The field of view of our lens is only \degrees(45) by \degrees(34), which is not enough to meet the competition field of view requirements of \degrees(60) vision in each direction.
+To address this issue, we designed a pan tilt unit which makes it possible to image the entire hemisphere below the fuselage of the aircraft.
 
-We took advantage of the MHS-PM5's swiveling lens assembly 
-(see Figure \ref{fig:camera_frontview}, and added a servo actuator which could rotate the lens \degrees{135} back from straight forward, as shown in Figure \ref{fig:pantilt_linkdetail}.
-The tilt servo is a Hitech HS 77BB \ref{hitechs77bb}, and is capable of \degrees{180} rotation.
+We took advantage of the MHS-PM5's swiveling lens assembly (see Figure \ref{fig:camera_frontview}, and added a servo actuator which can rotate the lens \degrees{135} back from straight forward, as shown in Figure \ref{fig:pantilt_linkdetail}.
+The tilt servo is a Hitech HS 77BB \cite{hitechs77bb}, and is capable of \degrees{180} rotation.
 Figures \ref{fig:pantilt} and \ref{fig:pantilt_linkdetail} show the camera lens tilted at \degrees{45} below the horizon.
@@
 A thrust bearing takes lateral loads off the pan servo shaft, and stabilizes the assembly.
-The pan servo is a Hitech HS 785HB \ref{hitechhs785}, and is capable of \degrees{1260} rotation (3.5 revolutions).
+The pan servo is a Hitech HS 785HB \cite{hitechs785}, and is capable of \degrees{1260} rotation (3.5 revolutions).
 Both servos are powered and controlled by the Pan-tilt control board.
 
-The total height of the imaging system is 6.75", a maximum 2'' of which is exposed below the fuselage of the aircraft.
-The rotating assembly is, on its widest dimension 2.13", and would fit inside a 2.5" cylinder.
-Wires exiting the side of the camera, inside the fuselage, may make the assembly a bit wider.
+The total height of the imaging system is 6.75", a maximum 2" of which is exposed below the fuselage of the aircraft.
+The rotating assembly is, on its widest dimension 2.13", and fits inside a 2.5" cylinder.
+Wires exiting the side of the camera, inside the fuselage, make the assembly a bit wider.
 Because the maximum rotation is only \degrees{1260}, the camera and tilt servo wires only have to be capable of wrapping \degrees{\pm 630} around the shaft below the thrust bearing.
-A wire guide will be added to keep the wires from tangling once construction of the pan-tilt is complete and installed in the aircraft.
+A wire guide hass been added to keep the wires from tangling.
 
@@
 Each picture taken by the camera is about 500KB in size, while our theoretical maximum throughput is only 115200 baud (and only about half of that in practice).
-Each full image takes about a minute to download, so we created a method to speed up the process.
-The entire image is first downloaded as an 800x600 resolution `thumbnail.'  Once the thumbnail is available for viewing, the operator can select an area of interest and download a high-res crop of just that area.
-Downloading both the crop and the thumbnail generally takes under 30 seconds, less than half the time of downloading the full image, allowing us to quickly obtain high resolution images of the targets of interest.
+Each full image takes almost three minutes to download, so we created a method to speed up the process.
+The entire image is first downloaded as an 800x600 low resolution `thumbnail.'  Once the thumbnail is available for viewing, the operator can select an area of interest and download a high-res crop of just that area.
+Downloading both the crop and the thumbnail generally takes under 30 seconds, less than $1/5$ the time of downloading the full image, allowing us to quickly obtain high resolution images of the targets of interest.
 
+\begin{figure} [H]
+	\centering
 		\begin{tabular}{rrr}
 		 & {\bf Size (KB)} & {\bf Download Time (s)} \\
- {\bf Original Image} &        500 &       69.4 \\
+		 {\bf Original Image} &        1200 &       166.7 \\
 		 {\bf Thumbnail} &        125 &       17.4 \\
@@
 		\end{tabular}
+	\caption[Download Times for Images] {size and download time for a full size image vs. a low-resolution thumbnail and a cropped area}
+	\label{fig:imagedownloadtimes}
+\end{figure}
 
@@
 \item Operator issues command to make images available for download
-\item Plane acknowledges request and begins transfer of images from internal camera memory to the BeagleBoard.  Sends notification when transfer is complete.
+\item Plane acknowledges request and begins transfer of images from internal camera memory to the BeagleBoard.  Plane sends notification when transfer is complete.
 \item Newly available image(s) are automatically put on a queue to be downloaded.
@@
 \item Steps 6-7 are repeated throughout the duration of the flight.
-These image downloading commands are low priority and will only be sent if there are no other commands to be executed, This allows for high priority commands like taking a picture or moving the camera to be executed at any time.
+These image downloading commands are low priority and will only be sent if there are no other commands to be executed.
+This allows for high priority commands like taking a picture or moving the camera to be executed at any time.
 \end{enumerate}
@@
 When an image has been fully downloaded to the ground targets can be identified and tagged with metadata.
-Information about the plane's orientation, gps coordinates, altitude, yaw/pitch/roll, and pan/tilt of the camera are sent down with the image.
+Information about the plane's orientation, gps coordinates, altitude, yaw/pitch/roll, and pan/tilt of the camera are sent down when the image is downloaded.
 Using this information the GPS location of the target can be determined using the intrinsic properties of the camera and some basic trigonometry.
@@
 \appendix
-\renewcommand\thesection{Appendix \Alph{section}}
+\renewcommand\thesection{Appendix \Alph{section}:}
 \section{Preflight Procedures}
@@
 \bibliographystyle{plain}
-\bibliography{paper}
+%\renewcommand{\refname}{\section{Citations}}
+\bibliography{citations}
 

### 2010-05-25 paper/paper.tex
changed documentclass to article. added chris eivan. made intro picture bigger. finished safety features. added link to source on github

@@
-\documentclass[10pt]{report}
+\documentclass[pdftex,10pt,letter]{article}
 \usepackage[pdftex]{graphicx}
@@
 
-\section{Introduction TWO}
+\section{Introduction}
 
@@
  Bill Perkowski,
+Chris Eivan,
 and  Mike Korb, 
-%and Chris (test pilot) 
 amongst many who spent time mentoring us.
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{../images/wreckage.jpg}
+	\includegraphics[width=0.75\textwidth]{../images/wreckage.jpg}
 	\caption{The recovery of our 2009 entry, the \emph{Icarus}}
@@
 
-We built a prototype using a Canon Powershot S5 IS \cite{powershotS5}, a camera which has no electrical remote control input.
+We built a prototype using a Canon Powershot S3 IS \cite{powershotS3}, a camera which has no electrical remote control input.
 We soldered wires to the push-button pads for the shutter and zoom buttons so we could override them remotely.
@@
 	\setlength{\parskip}{0cm}
-	\item Channel 5 (?) allows quick switch to safety pilot in case of autopilot malfunction.
-	\item If the plane loses contact with the ground station for more than XX seconds, the autopilot will default to a ``return home mode'' that will direct the plane towards the ground station in hopes of reestablishing a connection.
-If the connection remains dead for XX minutes, the autopilot will direct the plane to [DO SOMETHING?]
-	\item Channel 6 (?) functions as a remote kill switch, ensuring the ability to cut power to the engine if an emergent situation precludes direct manual shut off.
-	\item etc?
+	\item The channel 5 toggle switch determines whether the radio control transmitter, or the autopilot, has control over the flight control servos.
+The radio control safety pilot may disengage the autopilot at any time by flicking this switch.
+	\item If the plane loses telemetry with the ground station for more than 10 seconds, the autopilot will default to a ``return home mode'' that will direct the plane towards the ground station in hopes of reestablishing a connection.
+	\item If the connection telemetry remains dead for 30 seconds, the autopilot will bring the servos to a predefined default (throttle closed, elevator up, rudder right, ailerons left).
+The servos will remain in that position until the safety pilot takes control of the aircraft.
+	\item If the flight computer does not provide the servo controller with a new servo output value for longer than 2 seconds, the servo controller will bring the servos to a predefined default (throttle closed, elevator up, rudder right, ailerons left).
+The servos will remain in that position until the safety pilot takes control of the aircraft.
+	\item During engine start, the centerman (see \ref{appendix_preflight}) may kill the engine at any time with the ignition switch located on the right side of the aircraft, behind the engine mount.
+	\item Radio control channel 6 functions as a remote kill switch on the \emph{Knight Two}, disengaging the electronic ignition system of the engine.
+(The \emph{Daedalus} has a magneto ignition, so we only have the manual kill switch behind the engine mount)
 \end{itemize}
@@
 
+\newpage
+
 \bibliographystyle{plain}
 %\renewcommand{\refname}{\section{Citations}}
-\bibliography{citations}
+\bibliography{paper}
+
+The \LaTeX$\,$ source and figures of this paper are available on github:
+
+\url{http://github.com/pchickey/ruaiaa-2010doc/}
 

### 2010-05-25 paper/paper.tex
fiddled with some figure placements and some captions. started work on the autopilot description

@@
 \input{./title.tex}
-
-% We are definitely going to hit the page limit. No more abstract.
-%\begin{abstract}
-%We will write the abstract after the rest of the paper is finished. 
-%If we're really close to the page limit, we may be able to omit it.
-%\end{abstract}
 
@@
 Jeff Perkins, for his prolific networking and promotion of the team;
-Gregg Peters of BP Hobbies for getting us out of numerous binds;
-Dave Devries and Michael Maia, of Invensense, for their technical guidance;
+Gregg Peters, of BP Hobbies, for getting us out of numerous binds;
+Joe Vanderveer, for keeping a roof over our heads;
+Dave Devries and Michael Maia, of Invensense, for their technical guidance and donation of a sensor;
+F. Pasolini of ST Micro, for donation of a sensor;
 Professor Chung-chieh Shan, for overseeing Pat's independent study on the IMU design;
@@
 
+\newpage
 \subsection{Airframe}
@@
 Once rebuilt and covered,  the \emph{Daedalus} weighed 19.5 lbs with engine.
-A full 24oz fuel tank, plus 3.5 lbs electronics payload, would bring the maximum takeoff weight to 25.5lbs.
+A full 24oz fuel tank, plus 4lbs electronics payload, would bring the maximum takeoff weight to 25lbs.
 
@@
 The plane was carefully supported under the center section, nose, and tail while pre-weighed bags of sand were slowly placed symmetrically and simultaneously on the underside of each wing, starting from the innermost rib.
-The \emph{Daedalus} wings supported 64 lbs (250\% of the maximum take-off weight) under static loading.
+The \emph{Daedalus}'s wings supported 64 lbs under static loading.
+This is 260\% of the maximum take off weight, the force of a 1.5$g$ accelleration multiplied by a 1.73 safety factor.
 We also rocked the plane ten degrees from side to side while fully loaded to simulate differential loads during flight.
@@
 
-\begin{wrapfigure}{l}{0.4\textwidth}
+\subsubsection{Knight Two}
+\begin{wrapfigure}{r}{0.4\textwidth}
 	\centering
@@
 \end{wrapfigure}
-Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations
-
-\subsubsection{Knight Two}
 
@@
 
+\newpage
 \section{Autopilot}
@@
 
-\subsection{Requirements}
+\subsection{Hardware}
+\subsubsection{Single Board Computer}
+Beagleboard.
 
-We needed to interface with the following hardware components:
+\subsubsection{Other Components}
 \begin{itemize}
@@
 	\setlength{\parskip}{0cm}
-	\item GPS
-	\item servos
-	\item camera
-	\item etc.
+	\item GPS: 
+	\item Servo Controller:
+	\item Autopilot Radio:
+	\item Imaging Radio:
+	\item Wifi Adapter:
+	\item 
 \end{itemize}
 
-Interfacing with such a large array of devices meant we had to use a ...
-
-\subsection{Hardware}
-Description of components selected.
-Description of each input, output interface.
 \begin{figure}
@@
 \label{sec:autopilot_software}
+didn't want to be your ghost, didn't want to be anyone's ghost, didn't want to be your ghost, didn't want to be anyone's ghost
 
+but i don't want anyone else
 \begin{figure}
@@
 \subsection{Ground Station}
+
+\begin{wrapfigure}{r}{0.4\textwidth}
+	\includegraphics[width=0.37\textwidth]{../images/paparazzi_gcs.jpg}
+	\caption{Paparazzi GCS. Image Credit: Paparazzi Wiki\cite{}}
+	\label{fig:paparazzi_gcs}
+\end{wrapfigure}
 Because we based our autopilot software on the Paparazzi project, we were able to make use of their excellent ground control software (GCS) with essentially no modifications.
-
-Detail the capabilities of the Paparazzi GCS.
-
-Figure goes here: Paparazzi ground station screenshot, or whatever. 
-\subsection{Tuning and Testing}
-PAT
+The Paparazzi GCS provides a map interface that allows the operator and competition judges to inspect the aircraft's position to ensure it is within the allowed airspace.
+It also provides attitude and altitude telemetry, and notification of lost radio contact, so the operator can alert the safety pilot to unsafe operation.
 
@@
 		\end{tabular}
-		\caption[Pixel Width Mapping]{mapping of altitude to the area each pixel will represent in an image}
+		\caption[Pixel Width Mapping]{Linear resolution per pixel as a function of altitude, at directly below flight path and \degrees{45} off flight path}
 		\label{fig:pixelmapping}
@@
 In order to meet the target imaging requirments we needed to add both pan and tilt actuators to the camera.
-The field of view of our lens is only \degrees(45) by \degrees(34), which is not enough to meet the competition field of view requirements of \degrees(60) vision in each direction.
+The field of view of our lens is only \degrees{45} by \degrees{34}, which is not enough to meet the competition field of view requirements of \degrees{60} vision in each direction.
 To address this issue, we designed a pan tilt unit which makes it possible to image the entire hemisphere below the fuselage of the aircraft.
@@
 The endpoints of the tilt rotation rang are calibrated and enforced in software, because pushing the camera outside this range may damage it.
-
+\newpage
 \subsection{Obtaining Images}
@@
 		\end{tabular}
-	\caption[Download Times for Images] {size and download time for a full size image vs. a low-resolution thumbnail and a cropped area}
+	\caption[Download Times for Images] {Size and download time for a full size image vs. a low-resolution thumbnail and a cropped area}
 	\label{fig:imagedownloadtimes}

### 2010-05-25 paper/paper.tex
table headings

@@
 		\begin{tabular}{ccc}
-		\\{\bf altitude (ft)} & {\bf pixel width at 0$^{\circ}$ (in)} & {\bf pixel width at 45$^{\circ}$ (in)} \\
+		\\{\bf Altitude (ft)} & {\bf Pixel width at 0$^{\circ}$ (in)} & {\bf Pixel width at 45$^{\circ}$ (in)} \\
 	       100 &       0.38 &       0.54 \\

### 2010-05-25 paper/paper.tex
proofread from page 6 on, not including Autopilot specs or safety.

--- 2010-05-25 paper/paper.tex
@@
-\documentclass[pdftex,10pt,letter]{article}
+\documentclass[10pt]{report}
 \usepackage[pdftex]{graphicx}
@@
 
-\section{Introduction}
+\section{Introduction TWO}
 
-The Rutgers Autonomous Aircraft Team was formed in September 2008 with the goal of entering the 2009 Association for Unmanned Vehicle Systems International (AUVSI)\cite{auvsiweb} Unmanned Air Systems (UAS) competition\cite{uasweb}.
-We didn't make it to the 2009 competition after losing our aircraft only four weeks before.
+The Rutgers Autonomous Aircraft Team was formed in September 2008 with the goal of entering the 2009 AUVSI UAS competition.
+We didn't make it to the 2009 competition after loosing our aircraft only four weeks before.
 Our imaging system and autopilot system were also far from ready.
-In terms of NCAA sport eligibility, one could consider 2009 our `red shirt' year.
-In 2010 we are still rookies.
+
+In terms of NCAA sport eligibility, one could consider 2009 our `red shirt' year. In 2010 we are still rookies.
 
@@
  Bill Perkowski,
-Chris Eivan,
 and  Mike Korb, 
+%and Chris (test pilot) 
 amongst many who spent time mentoring us.
@@
 Rutgers Engineering Governance Council,
-Rutgers Alumni Association,
-WINLAB, a laboratory of the Rutgers University Department of Electrical Engineering, 
-and BP Hobbies.
+Rutgers Alumni Association, and
+WINLAB, a laboratory of the Rutgers University Department of Electrical Engineering.
+% What was Chris Brocco's Uncle's Company? Something Automation?
+BP Hobbies,
 
 We would like to acknowledge
-Louis Stumpf, for donating the aircraft that became the \emph{Daedalus};
+Louis Stumpf, for donating the aircraft which became the \emph{Daedalus};
 Jeff Perkins, for his prolific networking and promotion of the team;
@@
 Professor Chung-chieh Shan, for overseeing Pat's independent study on the IMU design;
-Ivan Seskar, for keeping us (and many other undergrads) in the black; and
+Ivan Seskar, for keeping us (and many other undergrads) in the black;
 Roger Kondos and Brij Pathak, for contributing to the thermopile attitude sensor.
@@
 	\centering
-	\includegraphics[width=0.75\textwidth]{../images/wreckage.jpg}
+	\includegraphics[width=0.6\textwidth]{../images/wreckage.jpg}
 	\caption{The recovery of our 2009 entry, the \emph{Icarus}}
@@
 \subsection{Goals}
-For this entry we took a close look at our goals and requirements, and attempted to meet the ``threshold'' competition requirements successfully.
+For this entry, we took a close look at our goals and requirements, and dialed back our approach to ensure we would meet the most basic competition requirements successfully.
 We started the design by forming, as a team, some broad goals based on our experience in the 2009 competition.
-These goals allowed sub-teams and individuals to tackle the smaller problems while still keeping an eye on the big picture.
+These formed the basis on which sub-teams and individuals tackled the smaller problems while still keeping an eye on the big picture.
 
-We also wanted to balance building components with buying components, keeping in mind our slim finances.
-While we couldn't build every part of our entry from scratch, we were motivated to build as much of our systems as possible for the learning experience.
-We tried to buy only components which we didn't have the time or resources to build ourselves.
+We also wanted to balance building with buying, keeping in mind our slim finances.
+While we couldn't build every part of our entry from scratch, we tried to build whatever we could given the 16th June competition deadline.
 
@@
 	\item The imaging system must be simple and capable.
-We assume the operator will be burdened with guiding the camera and identifying targets manually.
-We couldn't guarantee time and resources would permit us to add sensor information and control loops for camera guidance and computer vision for target identification.
+We assume the operator will be burdened with guiding the camera and identifying targets until time and resources permit us to add sensor information and control loops for guidance, and computer vision for identification.
+It will be custom built, because we could not find any affordable systems available commercially.
 \end{itemize}
 
-We will be performing a manual take off and giving the autopilot system control of the aircraft once the plane is in the air.
+We made design decisions of which competition goals we will attempt. 
+We will be performing a manual take off, and giving the autopilot system control of the aircraft once the plane is in the air.
 We will also perform a manual landing.
-We decided not to pursue the reach goals of autonomous takeoff and landing based on our inexperience and the risk of a crash during development.
+We decided not to pursue the reach goals of autonomous takeoff and landing based on our inexperience and the risk, during development and competition, to our aircraft.
 \subsection{System Design Decisions}
@@
 
+The teams made design decisions based on the aforementioned competition goals.
+
 One of the most difficult design decisions was to quantify ``ample payload.'' Payload size and weight is the major determining factor in aircraft size.
-We originally assumed a 2.5lb electronics (flight computer, sensors, and batteries) payload and a 2.5lb pan-tilt camera 8'' in diameter, requiring 6'' of vertical space below the bottom of the aircraft.
+We originally assumed a 2.5lb electronics (computer + batteries) payload and a 2.5lb pan-tilt camera 8'' in diameter, requiring 6'' of vertical space extruding from the bottom of the aircraft.
 These considerations introduced two major design parameters: the aircraft must be capable of carrying a minimum 5lb payload, and must be situated to provide clearance for the pan tilt unit.
 
-We also needed an easily controllable, stable platform to accommodate the autopilot, and we needed to keep flight speed low to make image capture as easy as possible.
-After consulting with R/C experts, a high wing trainer is looked like the most stable and proven fixed-wing design.
-Trainers traditionally have light wing loading for slow speed flight, and are able to carry a comparatively large payload.Their lower wing loading is easier to control, and easier to recover if power is lost.
+Additionally, we sought an easily controllable, stable platform to accommodate the autopilot, and an ability to fly at low speeds, in order to more easily capture clear images.
 
-Erring on the side of caution, we decided our aircraft should be capable of accommodating a worst-case scenario payload. 
-We were, in part, cautious due to a previous error where we rebuilt an entire aircraft to accommodate a larger payload.
-We decided on a high wing trainer with a 10' to 12' wingspan, so that we could accommodate up to 5lbs of payload while maintaining suitable wing loading.
-Though we managed to keep our payload weight well under 5lbs, the only major drawback to having an oversize aircraft is the difficulty of transport.
+We determined, after consulting with R/C experts, that a high wing trainer is the most stable and proven fixed-wing design.
+Trainers traditionally have light wing loading for slow speed flight, and are able to carry a comparatively large payload.
+Additionally, a large plane with lower wing loading is easier to control, and easier to recover if power is lost.
+Consciously agreeing to err on the side of caution, we sought a large plane capable of accommodating a worst-case scenario payload, and settled on a wingspan on the order of 10'-12'.
+Though lighter and smaller payload alternatives were found as the year progressed, these merely leave us with more freedom for future modifications.
+%That last sentence sounds kind of eh, I'm stuck on a decent wording, so I'm leaving it for now.
 
-Our selection of a trainer type aircraft encouraged us to rethink our pan tilt design.
-We based our initial assumptions of an 8'' diameter pan and tilt unit, mounted underneath the aircraft, based on a design that was to be mounted on the nose of our 2009 entry, an aircraft with a pusher motor configuration.
-When we considered the possibilities of either dramatically increasing the ride height of our aircraft, or making significant modifications to the width of the fuselage, we decided to redesign our pan tilt system to fit in a more suitable envelope.
+The decision to employ a large, high wing trainer, however, forced us to rethink our original pan-tilt design.
+Last year, in order to accommodate the pan-tilt unit, we designed a pusher plane, enabling us to mount the pan-tilt under the plane's nose, which had sufficient clearance.
+Since trainers are generally designed with front-mounted engines (I presume?), we were forced to engineer a new solution requiring significantly less clearance.
+In the long run, we modified the pan-tilt mount to allow us to recess more of the pan-tilt unit into the body of the plane---this had the added benefit of providing more protection to an expensive piece of equipment on takeoff and landing.
 
@@
 
-We had sketched plans for an electronics system built entirely on a Linux-based flight computer as early as July 2009, but planned on using a more standard microcontroller-based autopilot system with an auxiliary Linux-based computer for the imaging system until December 2009.
+We had sketched plans for an electronics system built entirely on a Linux-based flight computer as early as July 2009, but planned on using a more standard microcontroller-based autopilot system with an auxillary Linux-based computer for the imaging system until December 2009.
 After burning out our microcontroller based autopilot for the second time, we were faced with a long lead time for an expensive replacement part.
-On that impetus, we redesigned our electronics systems to eliminate the microcontroller autopilot and ensure all of our replacement parts would be inexpensive and commercially available from many vendors.
+On that impetus, we redesigned our electronics systems to eliminate the microcontroller autopilot and ensure all of our replacment parts would be inexpensive and commercially available from many vendors.
 The move to a single Linux-based flight computer, while not without obstacles, was a successful design decision, as we'll describe in this paper.
@@
 
-We felt the Paparazzi system was full featured and proven, but we weren't satisfied with the capabilities of the existing Paparazzi hardware.
+We felt the Paparazzi system was full featured and proven, but we weren't satisified with the capabilities of the existing Paparazzi hardware.
 The ``Tiny'' v2.1\cite{paparazzi_tinyv21}, the latest autopilot hardware released by the Paparazzi project, is based on a 64 pin ARM microcontroller.
@@
 Both Pat Hickey and Bradley Lord became intimately familiar with the source code and control flow inside the autopilot software, and all patches were reviewed and tested by both engineers.
-Numerous other safety measures are described in Section \ref{sec:autopilot_software}.
+Numerous other safety measures will be described in Section \ref{sec:autopilot_software}.
 
-Because our entry is essentially an experimental system and will always be operated with a safety pilot present, we engineered the safety pilot override to operate independently from the flight computer.
+Because our entry is essentially an experimental system, and will always be operated with a safety pilot present, we engineered the safety pilot override to operate independently from the flight computer.
 We have ensured that, short of losing power to the radio control receiver and servos, the safety pilot will always be able to take over with the flick of a single switch on his radio control transmitter.
@@
 
-When designing the imaging system our major goals were to have 
-1) a camera able to capture imagery ``up to \degrees{60} in all directions from vertically below the aircraft''\cite{competition_rules}, as per the rules; 
-2) a real-time video feed;
-4) the ability to download pictures while in flight; and
-3) the ability to remotely control the camera, particularly to trigger pictures.
-
-It was difficult to find a camera to base our imaging system on which met all of these requirements.
-Many still cameras can be fitted on a pan tilt system to requirement 1, and provide a real-time ``viewfinder'' video feed to meet requirement 2.
-Requirement 3 could typically be managed by attaching the camera to a single board computer by USB cable, though usually the camera powers down when attached to the computer, and some sort of switch is needed to deactivate the USB cable once the data transfer is complete.
-
-Requirement 4 had the greatest effect on our design decision.
-Typically, only high-end SLR-type still cameras provide a socket for remote control.
-These are beyond our price range, and are much larger and heavier than we would like.
-
-We built a prototype using a Canon Powershot S3 IS \cite{powershotS3}, a camera which has no electrical remote control input.
-We soldered wires to the push-button pads for the shutter and zoom buttons so we could override them remotely.
-Those wires were attached to isolating transistors, and a microcontroller was used for control.
-After some time, this clever circumvention of the factory anti-static shielding resulted in a burnt out camera.
-We abandoned this scheme.
+When designing the imaging system our major goals were to have 1) a camera with XXX range of motion.
+2) real-time video feed from the plane.
+3) ability to control the camera in real-time and take pictures.
+4) ability to download pictures while in flight.
+For the camera we chose the Sony Bloggie \cite{bloggie} because it was small, light, cheap, and took high resolution images and video.
+In addition it came with a built in rotating lens, which simplified the construcion of the pan-tilt unit, and it can also be controlled remotely over usb, so it interfaced with our flight linux computer.
+To communicate with the plane we developed custom ground station software to meet our needs and installed two sets of radios, one for the video feed, and another for communicating with the plane and downloading images.
+These radios are independent from the set of radios used for the autopilot because we didn't want to have dependencies between the two systems.
 
@@
 At the time of this writing, we have completed and flown a single flight vehicle, which we have named the \emph{Daedalus}.
-We have also nearly completed a very similar backup airframe which will serve as a backup in case the \emph{Daedalus} has a disaster.
+We have also nearly completed a very similar backup airframe which serve as a backup in case the \emph{Daedalus} has a disaster.
 The \emph{Knight Two} will be fitted with nearly identical R/C flight electronics, and accept the same autopilot and imaging system payload as the \emph{Daedalus}.
@@
 \subsection{Airframe}
+
+We elected to pursue a kit-type airplane rather than our own design. 
+Our own custom design, the \emph{Icarus}, suffered a catastrophic structural failure last year.
+From this, we learned that custom structures require much more testing and revision than a proven kit.
+We also reconsidered our payload requirements and decided we could easily modify a kit plane to accept our autopilot and imaging system.
 
@@
 \end{wrapfigure}
-Louis Stumpf, a member of our local R/C club, donated a 10ft 1in (3.07m) wingspan, 7ft 8in (2.33m) long, high wing trainer to our club. It met all of the requirements outlined previously.
-
+A member of our local R/C club, Tri-County RC \cite{tricountyRC}, donated a ten-foot wingspan high wing trainer to our club.
 The plane, first built out of foam and balsa from long-lost plans in 1986, was accepted graciously, but required quite a bit of work before it was once again flight-worthy.
@@
 We stripped and rebuilt nearly the entire airframe, adding carbon-fiber reinforcements to the wing in anticipation of increased wing loading due to our payload.
-Once rebuilt and covered,  the \emph{Daedalus} weighed 19.5 lbs with engine.
+Once rebuilt and covered, we tested the \emph{Daedalus} weighed 19.5 lbs with engine.
 A full 24oz fuel tank, plus 3.5 lbs electronics payload, would bring the maximum takeoff weight to 25.5lbs.
@@
 We wanted to certify the wings and wing root would withstand large dynamic forces in flight. 
-We designed a test to mimic the 
+We designed a test similar to mimic the 
 Federal Airworthiness Standards' structural requirements\footnote{Special Federal Aviation Regulation No.
@@
 
-Satisfied, we made engine and surface tests (see \ref{sec:appendix_preflight}) before taxiing the plane around a parking lot.
+Satisfied, made engine and surface tests (see \ref{sec:appendix_preflight}) before taxiing the plane around a parking lot.
 We brought the plane back inside, examined it for signs of stress and, finding everything in good condition, deemed the \emph{Daedalus} flight-worthy.
@@
 We selected an Aero-Design 12 Foot Telemaster \cite{aerodesign} because it was available as a balsa kit and similar in size to the \emph{Daedalus}.
-The 12 Foot Telemaster has a 12ft 1in (3.68m) wingspan, and is 7ft 10in (2.38m) in length.We named it the \emph{Knight Two} after the Rutgers University mascot, the Scarlet Knight.
 
-At the time of this writing, the \emph{Knight Two} is 85\% complete. We plan to complete the aircraft and bring it to the competition as a spare.  
-Since our testing plan for the \emph{Daedalus} was successful, we will follow the same plan when the \emph{Knight Two} is completed.
+We named it the \emph{Knight Two} after the Rutgers University mascot, the Scarlet Knight.
+We may rename it if someone can think of a better name.
+Pat needed to think up a decent name real fast in order to make this rough draft.
+
+Since our testing plan for the \emph{Daedalus} was successful, we will follow the same plan when, in the coming weeks, we finish building \emph{Knight Two}.
+
+Someone please detail the outer dimensions. The inner dimensions, space used for flight electronics, space used for payload underneath. Landing gear height. Number of servos and their locations
+
+Figure may go here: Line drawing of Knight Two. JOHN
 
@@
 	\setlength{\parskip}{0cm}
-	\item The channel 5 toggle switch determines whether the radio control transmitter, or the autopilot, has control over the flight control servos.
-The radio control safety pilot may disengage the autopilot at any time by flicking this switch.
-	\item If the plane loses telemetry with the ground station for more than 10 seconds, the autopilot will default to a ``return home mode'' that will direct the plane towards the ground station in hopes of reestablishing a connection.
-	\item If the connection telemetry remains dead for 30 seconds, the autopilot will bring the servos to a predefined default (throttle closed, elevator up, rudder right, ailerons left).
-The servos will remain in that position until the safety pilot takes control of the aircraft.
-	\item If the flight computer does not provide the servo controller with a new servo output value for longer than 2 seconds, the servo controller will bring the servos to a predefined default (throttle closed, elevator up, rudder right, ailerons left).
-The servos will remain in that position until the safety pilot takes control of the aircraft.
-	\item During engine start, the centerman (see \ref{appendix_preflight}) may kill the engine at any time with the ignition switch located on the right side of the aircraft, behind the engine mount.
-	\item Radio control channel 6 functions as a remote kill switch on the \emph{Knight Two}, disengaging the electronic ignition system of the engine.
-(The \emph{Daedalus} has a magneto ignition, so we only have the manual kill switch behind the engine mount)
+	\item Channel 5 (?) allows quick switch to safety pilot in case of autopilot malfunction.
+	\item If the plane loses contact with the ground station for more than XX seconds, the autopilot will default to a ``return home mode'' that will direct the plane towards the ground station in hopes of reestablishing a connection.
+If the connection remains dead for XX minutes, the autopilot will direct the plane to [DO SOMETHING?]
+	\item Channel 6 (?) functions as a remote kill switch, ensuring the ability to cut power to the engine if an emergent situation precludes direct manual shut off.
+	\item etc?
 \end{itemize}
@@
 
-\newpage
-
 \bibliographystyle{plain}
 %\renewcommand{\refname}{\section{Citations}}
-\bibliography{paper}
-
-The \LaTeX$\,$ source and figures of this paper are available on github:
-
-\url{http://github.com/pchickey/ruaiaa-2010doc/}
+\bibliography{citations}
 

### 2010-05-25 paper/paper.tex
expanding autopilot section, moved figures a little

--- 2010-05-25 paper/paper.tex
@@
 \subsection{Hardware}
-\subsubsection{Single Board Computer}
-Beagleboard.
-
-\subsubsection{Other Components}
-\begin{itemize}
-	\setlength{\itemsep}{0cm}
-	\setlength{\parskip}{0cm}
-	\item GPS: 
-	\item Servo Controller:
-	\item Autopilot Radio:
-	\item Imaging Radio:
-	\item Wifi Adapter:
-	\item 
-\end{itemize}
-
 \begin{figure}
@@
 \end{figure}
+
+\subsubsection{Single Board Computer}
+We selected the Beagleboard \cite{beagleboard}, a single board computer made by Texas Instruments.
+It features a 500MHz ARM Cortex A8 processor and 256MB ram.
+We installed Angstrom Linux \cite{angstromlinux} on an 8GB SD card.
+The Beagleboard has a few logic level serial, parallel, and general purpose IO interfaces on board, but for simplicity we chose to attach all of the devices through the USB bus.
+The Beagleboard provides one USB host port for this purpose, but the power supply can only sustain a low current draw.
+To provide more current to the attached devices, we use a Belkin 7-port powered USB hub \cite{belkin7usbhub}.
+With a proper power supply, the USB hub will provide up to 500mA to each port on the 5v/GND circuit.
+
+\subsubsection{Servo Subsystem}
+We are using Pololu Micro Maestro \cite{micromaestro}, which communicates with the flight computer over USB and features 6 Servo PWM outputs.
+On Linux, the Micro Maestro enumerates as a \texttt{ttyACM} device.
+A simple serial protocol is used to update the servo PWM outputs.
+The Micro Maestro also features a configurable timeout and failsafe position for the servo PWM outputs.
+We have configured the device to bring each servo to a predefined positon (throttle closed, elevator up, rudder right, ailerons left) after a 2 second new command timeout.
+This ensures that if the autopilot process or entire flight computer fails to respond for 2 seconds, the aircraft will enter this mode until the safety pilot takes control.
+
+The safety pilot is able to take control of the aircraft
+
+\subsubsection{Other Components}
+\begin{itemize}
+	\setlength{\itemsep}{0cm}
+	\setlength{\parskip}{0cm}
+	\item GPS: We selected the GS407 \cite{gs407gps}, a uBlox LEA-5H \cite{ublox_lea5h} based GPS with an active helical antenna.
+We picked the unit because the uBlox chipset has been used successfully with Paparazzi in the past, and provides an internal Kalman filter mode for aircraft expecting accellerations of $\pm2g$.
+The GPS is powered and connected to the USB hub with a Sparkfun FTDI adapter \cite{sparkfun_ftdi}.
+	\item Autopilot Radio:
+	\item Imaging Radio:
+	\item Wifi Adapter:
+\end{itemize}
+\subsubsection{Power}
+Both the Beagleboard and the powered USB hub are provided with 5VDC via a switching regulator and a 3 cell (11.1v nominal) 1.5Ah Lithium Polymer battery.
+This provides, on average, 90 minutes of operation with all devices on.
 
@@
 	\includegraphics[width=0.9\textwidth]{../images/software_comm_diagram.pdf}
-	\caption{Communications diagram}	
+	\caption{Software communications diagram}	
 	\label{fig:software_comm_diagram}
@@
 	\includegraphics[width=0.37\textwidth]{../images/paparazzi_gcs.jpg}
-	\caption{Paparazzi GCS. Image Credit: Paparazzi Wiki\cite{}}
+	\caption{Paparazzi GCS. Image Credit: Paparazzi Wiki \cite{paparazziweb }}
 	\label{fig:paparazzi_gcs}

### 2010-05-25 paper/paper.tex
finished up other components.

@@
 
-The safety pilot is able to take control of the aircraft
+A Futaba 2.4GHz ``FAAST'' 6-channel radio transmitter and receiver \cite{futaba24} is the safety radio for our entry.
+We've tested the radio in the presence of our other 2.4GHz radios (Wifi adapter, autopilot radio, and imaging system radio) at a variety of ranges and configurations, and are confident that any radio interference is not harmful to the response of the servo outputs.
+
+A Pololu 4 channel servo multiplexer \cite{pololumux} allows safety pilot is able to select whether the signals from the Futaba radio control receiver or the autopilot are in control of the servos.
+The four control channels (throttle, elevator, rudder, ailerons) are routed from both the Futaba receiver and Micro Maestro to the Master and Slave inputs, respectively, on the multiplexer.
+The multiplexer and servos are powered by the Futaba receiver, which has an independent power supply from the flight computer.
+As long as the Futaba receiver has power and signal from the radio control transmitter, the safety pilot will be able to take control of the aircraft.
+
+\subsubsection{Attitude and Heading Reference Sensor}
 
@@
 The GPS is powered and connected to the USB hub with a Sparkfun FTDI adapter \cite{sparkfun_ftdi}.
-	\item Autopilot Radio:
-	\item Imaging Radio:
-	\item Wifi Adapter:
+	\item Autopilot Radio: We selected a Maxstream XBee Pro (Series 2.5) 2.4GHz 50mW \cite{xbeepro24} radio system for the autopilot telemetry and uplink radio.
+It operates at the relatively slow baud rate of 19200bps.
+We selected this system because it is known to work well with the Paparazzi system, and it can provide a seamless point-to-point UART pipe with two transcievers.
+Importantly, it deals well with interference from other 2.4GHz radio connections nearby, including other XBee Pro radios, the Futaba ``FAAST'' radio control signal, and 2.4GHz 802.11g Wifi.
+	\item Imaging Radio: We are using an identical pair of XBee Pro radios in a point-to-point 115200bps UART connection configuration to connect the imaging system to the ground.
+We have tested and found that, when operating two point-to-point connections at once, neither suffers measurably at up to 1/3 mile line of sight range.
+Therefore, rather than write complicated software to use both the autopilot serial communication protocol and the imaging radio serial communication protocol with the same XBee radio, we use two radio connections on the same band with little penalty.
+	\item Wifi Adapter: We use a 802.11g (wifi) adapter for convenience in the lab and at the field to program the flight computer and supervise the execution of autopilot proceses.
+The reliability of our wifi connection drops off significantly beyond 600ft (200m), so we do not actively use the link during flight or for any telemetry.
 \end{itemize}

### 2010-05-25 paper/paper.tex
killed a fixed figure and made it smaller to free up space again
began ahrs

@@
 \subsubsection{Daedalus}
-\begin{wrapfigure}{l}{0.4\textwidth}
-	\includegraphics[width=0.37\textwidth]{../images/daedalus_isometric.jpg}
+\begin{wrapfigure}{l}{0.5\textwidth}
+	\includegraphics[width=0.47\textwidth]{../images/daedalus_isometric.jpg}
 	\caption{The \emph{Daedalus}}
@@
 
+Early in the school year, we built a thermopile-based attitude (pitch and roll) sensor very similar to the sensor design used by the Paparazzi Tiny autopilot.
+Accompanied by a driver which allowed it to appear to the airborne software just as it would on the Tiny, this simple analog sensor helped us prototype and test the Paparazzi system.
+Having built a superior sensor, we keep it as a backup.
+It is detailed in Figure \ref{fig:hardware_conn_diagram} although we do not intend to use it for the competition.
+
+Thanks to sensor donations from Invensense and ST Micro, we were able to build a 9 axis Attitude and Heading Reference Sensor (AHRS) from an Invensense ITG-3200 3-axis rate gyro \cite{itg3200} and an ST Micro LSM303DLH 3-axis accelerometer and magnetometer.
+\cite{lsm303dlh}
+
 \subsubsection{Other Components}
@@
 	\subfloat[Photograph of pan tilt unit installed in \emph{Daedalus}]{\label{fig:pantilt_photo}
-			\includegraphics[width=0.75\textwidth]{../images/pantilt_photo1.jpg}}
+			\includegraphics[width=0.55\textwidth]{../images/pantilt_photo1.jpg}}
 	\caption{Pan Tilt Unit}
@@
 
-\begin{figure} [H]
+\begin{figure}
   \centering
-  	\includegraphics[width=0.9\textwidth]{../images/ImageStationControls.jpg}
+  	\includegraphics[width=0.5\textwidth]{../images/ImageStationControls.jpg}
   	\caption[Image Station Control Panel]{The Control Panel for the Image Station.

### 2010-05-25 paper/paper.tex
made the side margins 1.25 from 1.5, and re-arranged a handful of figures to straighten things out. should have plenty of room now

@@
 \usepackage[pdftex]{graphicx}
-\usepackage[left=1.5in, right=1.5in, top=1.25in, bottom=1.25in]{geometry}
+\usepackage[left=1.25in, right=1.25in, top=1.25in, bottom=1.25in]{geometry}
 \usepackage{hyperref}
@@
 
-\newpage
 \subsection{Airframe}
@@
 \subsubsection{Daedalus}
-\begin{wrapfigure}{l}{0.5\textwidth}
-	\includegraphics[width=0.47\textwidth]{../images/daedalus_isometric.jpg}
-	\caption{The \emph{Daedalus}}
-	\label{fig:daedalus}
-\end{wrapfigure}
+
 Louis Stumpf, a member of our local R/C club, donated a 10ft 1in (3.07m) wingspan, 7ft 8in (2.33m) long, high wing trainer to our club. It met all of the requirements outlined previously.
@@
 A full 24oz fuel tank, plus 4lbs electronics payload, would bring the maximum takeoff weight to 25lbs.
-
+\begin{wrapfigure}{l}{0.5\textwidth}
+	\includegraphics[width=0.47\textwidth]{../images/daedalus_isometric.jpg}
+	\caption{The \emph{Daedalus}}
+	\label{fig:daedalus}
+\end{wrapfigure}
 We wanted to certify the wings and wing root would withstand large dynamic forces in flight. 
@@
 \subsubsection{Knight Two}
-\begin{wrapfigure}{r}{0.4\textwidth}
+\begin{figure}
 	\centering
@@
 			\includegraphics[width=0.37\textwidth]{../images/knighttwo_uncovered_outdoors.jpg}}
-\\
 	\subfloat[A completed 12 Foot Telemaster. Photo Credit: Hobby Lobby] {\label{fig:knighttwo_12ft_telemaster} 
@@
 
-\end{wrapfigure}
+\end{figure}
 
@@
 
-\newpage
 \section{Autopilot}
@@
 \subsection{Hardware}
-\begin{figure}
+\begin{figure}[t]
 	\centering
@@
 
-Thanks to sensor donations from Invensense and ST Micro, we were able to build a 9 axis Attitude and Heading Reference Sensor (AHRS) from an Invensense ITG-3200 3-axis rate gyro \cite{itg3200} and an ST Micro LSM303DLH 3-axis accelerometer and magnetometer.
-\cite{lsm303dlh}
+Thanks to sensor donations from Invensense and ST Micro, we were able to build a 9 axis Attitude and Heading Reference Sensor (AHRS) from an Invensense ITG-3200 3-axis rate gyro \cite{itg3200} and an ST Micro LSM303DLH 3-axis accelerometer and magnetometer \cite{lsm303dlh}.
 
@@
 The endpoints of the tilt rotation rang are calibrated and enforced in software, because pushing the camera outside this range may damage it.
-\newpage
+
 \subsection{Obtaining Images}

### 2010-05-25 paper/paper.tex
spel check and some additions to autopilot software

@@
 The \emph{Daedalus}'s wings supported 64 lbs under static loading.
-This is 260\% of the maximum take off weight, the force of a 1.5$g$ accelleration multiplied by a 1.73 safety factor.
+This is 260\% of the maximum take off weight, the force of a 1.5$g$ acceleration multiplied by a 1.73 safety factor.
 We also rocked the plane ten degrees from side to side while fully loaded to simulate differential loads during flight.
@@
 The Micro Maestro also features a configurable timeout and failsafe position for the servo PWM outputs.
-We have configured the device to bring each servo to a predefined positon (throttle closed, elevator up, rudder right, ailerons left) after a 2 second new command timeout.
+We have configured the device to bring each servo to a predefined position (throttle closed, elevator up, rudder right, ailerons left) after a 2 second new command timeout.
 This ensures that if the autopilot process or entire flight computer fails to respond for 2 seconds, the aircraft will enter this mode until the safety pilot takes control.
@@
 \subsubsection{Attitude and Heading Reference Sensor}
-
+\begin{wrapfigure}{l}{0.3\textwidth}
+	\centering
+	\rotatebox{-90}{
+		\includegraphics[width=0.37\textwidth]{../images/ahrs_sensor_top.jpg}
+	}
+	\caption{Completed AHRS sensor}
+	\label{fig:ahrs_picture}
+\end{wrapfigure}
 Early in the school year, we built a thermopile-based attitude (pitch and roll) sensor very similar to the sensor design used by the Paparazzi Tiny autopilot.
@@
 Thanks to sensor donations from Invensense and ST Micro, we were able to build a 9 axis Attitude and Heading Reference Sensor (AHRS) from an Invensense ITG-3200 3-axis rate gyro \cite{itg3200} and an ST Micro LSM303DLH 3-axis accelerometer and magnetometer \cite{lsm303dlh}.
+Both the ITG-3200 and LSM303DLH have an all digital interface, and communicate with a microcontroller on a single I2C \cite{i2c} serial bus.
+Samples of all 9 axes (three each of gyro, accelerometer, and magnetometer) are taken and sent to the flight computer at 40Hz.
+
+The two sensors and an AVR microcontroller, mounted on breakout boards, were hooked up using simple 0.1" pitch protoboard, as shown in Figure \ref{fig:ahrs_picture}.
+The sensor board is powered by the Sparkfun FTDI serial to USB adapter \cite{sparkfun_ftdi} the microcontroller uses to communicate with the flight computer.
 
@@
 	\item GPS: We selected the GS407 \cite{gs407gps}, a uBlox LEA-5H \cite{ublox_lea5h} based GPS with an active helical antenna.
-We picked the unit because the uBlox chipset has been used successfully with Paparazzi in the past, and provides an internal Kalman filter mode for aircraft expecting accellerations of $\pm2g$.
-The GPS is powered and connected to the USB hub with a Sparkfun FTDI adapter \cite{sparkfun_ftdi}.
+We picked the unit because the uBlox chipset has been used successfully with Paparazzi in the past, and provides an internal Kalman filter mode for aircraft expecting accelerations of $\pm2g$.
+The GPS is powered and connected to the USB hub with a Sparkfun FTDI serial to USB adapter \cite{sparkfun_ftdi}.
 	\item Autopilot Radio: We selected a Maxstream XBee Pro (Series 2.5) 2.4GHz 50mW \cite{xbeepro24} radio system for the autopilot telemetry and uplink radio.
 It operates at the relatively slow baud rate of 19200bps.
-We selected this system because it is known to work well with the Paparazzi system, and it can provide a seamless point-to-point UART pipe with two transcievers.
+We selected this system because it is known to work well with the Paparazzi system, and it can provide a seamless point-to-point UART pipe with two transceivers.
 Importantly, it deals well with interference from other 2.4GHz radio connections nearby, including other XBee Pro radios, the Futaba ``FAAST'' radio control signal, and 2.4GHz 802.11g Wifi.
@@
 Therefore, rather than write complicated software to use both the autopilot serial communication protocol and the imaging radio serial communication protocol with the same XBee radio, we use two radio connections on the same band with little penalty.
-	\item Wifi Adapter: We use a 802.11g (wifi) adapter for convenience in the lab and at the field to program the flight computer and supervise the execution of autopilot proceses.
+	\item Wifi Adapter: We use a 802.11g (wifi) adapter for convenience in the lab and at the field to program the flight computer and supervise the execution of autopilot process.
 The reliability of our wifi connection drops off significantly beyond 600ft (200m), so we do not actively use the link during flight or for any telemetry.
@@
 
-\subsection{Software}
+\subsection{Autopilot Software}
 \label{sec:autopilot_software}
-didn't want to be your ghost, didn't want to be anyone's ghost, didn't want to be your ghost, didn't want to be anyone's ghost
 
-but i don't want anyone else
+The autopilot software is a port of the Paparazzi airborne software to Linux. The airborne software is written in C and originally assumed a bare-metal (that is, no operating system)
+
 \begin{figure}
@@
 Because of the swiveling lens we were able to achieve two rotational degrees of freedom while only rotating the body of the camera about one axis.
-This greatly simiplified the pan-tilt design, and also allowed us to easily mount the camera primarily inside the fuselage of the aircraft, which protects it from damage during takeoff and landing, and reduces drag on the airframe.
+This greatly simplified the pan-tilt design, and also allowed us to easily mount the camera primarily inside the fuselage of the aircraft, which protects it from damage during takeoff and landing, and reduces drag on the airframe.
 
@@
 Figures \ref{fig:pantilt} and \ref{fig:pantilt_linkdetail} show the camera lens tilted at \degrees{45} below the horizon.
-The camera and tilt servo assembley is attached to the aircraft on a rotating mount.
+The camera and tilt servo assembly is attached to the aircraft on a rotating mount.
 A thrust bearing takes lateral loads off the pan servo shaft, and stabilizes the assembly.

### 2010-05-25 paper/paper.tex
finishing up autopilot software etc

@@
 Both the ITG-3200 and LSM303DLH have an all digital interface, and communicate with a microcontroller on a single I2C \cite{i2c} serial bus.
+The microcontroller performs the self-test on the LSM303DLH at start-up, and proceeds to send uncalibrated samples to the flight computer over its UART connection.
 Samples of all 9 axes (three each of gyro, accelerometer, and magnetometer) are taken and sent to the flight computer at 40Hz.
@@
 The sensor board is powered by the Sparkfun FTDI serial to USB adapter \cite{sparkfun_ftdi} the microcontroller uses to communicate with the flight computer.
-
 \subsubsection{Other Components}
@@
 \label{sec:autopilot_software}
-
-The autopilot software is a port of the Paparazzi airborne software to Linux. The airborne software is written in C and originally assumed a bare-metal (that is, no operating system)
-
-\begin{figure}
+\begin{figure}[t]
 	\centering
@@
 
+The autopilot software is a port of the Paparazzi airborne software to Linux.
+The airborne software is written in C and originally assumed a bare-metal (that is, no operating system) microcontroller environment with memory-mapped IO.
+The port mostly constituted inserting several layers between blocking Linux system calls and the control loop's expectation of non-blocking memory reads.
+The simplest way to do this was by splitting the autopilot process into separate threads for aircraft control and device access.
+We use mutexed shared memory to communicate between these threads, which works well because the aircraft control thread is a loop which only runs (therefore, only needs a lock) 60 times per second.
+
+The multithread approach helps ensure that the aircraft control thread will not be disrupted by busy waiting.
+It results in a simple structure of the whole program, and minimized the amount of code we had to change in the port.
+We care about minimizing changes to Paparazzi because the code is known to be well tested and reliable.
+
+The multithread approach works well for resources which are only needed by the autopilot process, but both the attitude and GPS information are needed by both the autopilot and imaging system processes.
+To remedy this, we use \texttt{gpsd} \cite{gpsd}, the standard Linux GPS service, to interface with the uBlox GPS device over a serial port.
+\texttt{Libgps}, a library for interfacing with \texttt{gpsd}, provides updates to client threads in both the autopilot and imaging processes.
+We handle the AHRS updates in the same way, except instead of the specialized \texttt{libgps} library we use the generic \texttt{mqueue} \cite{mqueue} service, a message passing facility provided by the Linux kernel.
+Each client thread opens a named queue, and the AHRS process delivers updated measurements to each queue.
+
+\subsection{Attitude and Heading Reference Sensor Software}
+
+The physical AHRS board sends 9 uncalibrated measurements to the flight computer, but the aircraft control loop is requires in attitude and heading expressed as 3 Euler angles \cite{euler_angles} and their derivatives with respect to time.
+It is straightforward to estimate Euler angle derivatives from rate gyro measurements, but estimating three degrees of angular orientation from six magnetometer and accelerometer measurements is a difficult nonlinear operation.
+Many algorithms exist
+
 \subsection{Ground Station}
@@
 \begin{wrapfigure}{r}{0.4\textwidth}
+	\vspace{-20pt}
+	\centering
 	\includegraphics[width=0.37\textwidth]{../images/paparazzi_gcs.jpg}
@@
 	\label{fig:paparazzi_gcs}
+	\vspace{-60pt}
 \end{wrapfigure}

### 2010-05-25 paper/paper.tex
all aboard

@@
 In order to mitigate a disaster such as we had last year, we decided to build another airframe which could replace the \emph{Daedalus} at short notice.
-We selected an Aero-Design 12 Foot Telemaster \cite{aerodesign} because it was available as a balsa kit and similar in size to the \emph{Daedalus}.
+We selected an Aero Craft Ltd.
+12 Foot Telemaster \cite{aero_craft_ltd} because it was available as a balsa kit and similar in size to the \emph{Daedalus}.
 The 12 Foot Telemaster has a 12ft 1in (3.68m) wingspan, and is 7ft 10in (2.38m) in length.We named it the \emph{Knight Two} after the Rutgers University mascot, the Scarlet Knight.
@@
 The Beagleboard provides one USB host port for this purpose, but the power supply can only sustain a low current draw.
-To provide more current to the attached devices, we use a Belkin 7-port powered USB hub \cite{belkin7usbhub}.
+To provide more current to the attached devices, we use a Belkin 7-port powered USB hub.
 With a proper power supply, the USB hub will provide up to 500mA to each port on the 5v/GND circuit.
@@
 Thanks to sensor donations from Invensense and ST Micro, we were able to build a 9 axis Attitude and Heading Reference Sensor (AHRS) from an Invensense ITG-3200 3-axis rate gyro \cite{itg3200} and an ST Micro LSM303DLH 3-axis accelerometer and magnetometer \cite{lsm303dlh}.
-Both the ITG-3200 and LSM303DLH have an all digital interface, and communicate with a microcontroller on a single I2C \cite{i2c} serial bus.
+Both the ITG-3200 and LSM303DLH have an all digital interface, and communicate with a microcontroller on a single I2C \cite{wikii2c} serial bus.
 The microcontroller performs the self-test on the LSM303DLH at start-up, and proceeds to send uncalibrated samples to the flight computer over its UART connection.
@@
 To remedy this, we use \texttt{gpsd} \cite{gpsd}, the standard Linux GPS service, to interface with the uBlox GPS device over a serial port.
-\texttt{Libgps}, a library for interfacing with \texttt{gpsd}, provides updates to client threads in both the autopilot and imaging processes.
+\texttt{libgps}, a library for interfacing with \texttt{gpsd}, provides updates to client threads in both the autopilot and imaging processes.
 We handle the AHRS updates in the same way, except instead of the specialized \texttt{libgps} library we use the generic \texttt{mqueue} \cite{mqueue} service, a message passing facility provided by the Linux kernel.
@@
 
-The physical AHRS board sends 9 uncalibrated measurements to the flight computer, but the aircraft control loop is requires in attitude and heading expressed as 3 Euler angles \cite{euler_angles} and their derivatives with respect to time.
+The physical AHRS board sends 9 uncalibrated measurements to the flight computer, but the aircraft control loop is requires in attitude and heading expressed as 3 Euler angles and their derivatives with respect to time.
 It is straightforward to estimate Euler angle derivatives from rate gyro measurements, but estimating three degrees of angular orientation from six magnetometer and accelerometer measurements is a difficult nonlinear operation.
-Many algorithms exist
+
+While their discussion is outside the scope of this paper, many algorithms exist for this conversion.
+We use a Quaternion Kalman Filter algorithm described in \cite{choukroun04}, implemented as a complete filter in Haskell.
+Angular orientation is tracked in quaternion form, and translated into euler angles for the use of the autopilot and imaging processes.
+This implementation was carried out as part of an independent study by team member Pat Hickey, and more details can be found in \cite{hickey10hsqkf}.
 
@@
 	\label{fig:paparazzi_gcs}
-	\vspace{-60pt}
+	\vspace{-40pt}
 \end{wrapfigure}
@@
 The servos will remain in that position until the safety pilot takes control of the aircraft.
-	\item During engine start, the centerman (see \ref{appendix_preflight}) may kill the engine at any time with the ignition switch located on the right side of the aircraft, behind the engine mount.
+	\item During engine start, the centerman (see \ref{sec:appendix_preflight}) may kill the engine at any time with the ignition switch located on the right side of the aircraft, behind the engine mount.
 	\item Radio control channel 6 functions as a remote kill switch on the \emph{Knight Two}, disengaging the electronic ignition system of the engine.
@@
 
-\newpage
-
 \bibliographystyle{plain}
